---
title: "04.1.4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup 

```{r}
"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")
library("GenomicRanges")
local.dir <- "/Users/jasont/science/projects/wtsa/promoter_driven/"
rescomp.dir <- "/Users/jasont/science/servers/FUSE5/"
work.dir <- rescomp.dir %&% "projects/wtsa/promoter_driven/"
input.dir <- work.dir %&% "peaky_interactions/"
output.dir <- work.dir %&% "peaky_interactions/track_files/"
```

```{r}
prom.df <- fread(input.dir %&% "promoter_endo_peaky-interactions.txt")
prom.full.df <- fread(input.dir %&% "promoter_endo_peaky-output.txt")
```

Note: For visualization purposes, need to remove duplicate fragments to generate bigwig files 
```{r}
library("GenomicRanges")
create_track_dfs <- function(peak.df,full.df){
  # baseMean	log2FoldChange	lfcSE	stat	pvalue	padj
  bed.df <- dplyr::select(peak.df,one_of("seqnames","start","end","capture"))
  full.sub.df <- dplyr::select(full.df,one_of("seqnames","start","end"))
  ##full.sub.df <- full.sub.df[!duplicated(full.sub.df$start),]
  full.gr <- GRanges(seqnames=full.df$seqnames,
                           IRanges(start=full.df$start,end=full.df$end))
  narrow.gr <- narrow(full.gr,4,-4)
  narrow.df <- as.data.frame(narrow.gr)
  narrow.df$seqnames <- as.character(narrow.df$seqnames)

  rjmcmc.df <- narrow.df[,1:3]
  rjmcmc.df$rjmcmc_pos <- full.df$rjmcmc_pos
  ##disjoin.df <- disjoin(narrow.gr) %>% as.data.frame() %>% dplyr::filter(width!=4)
  ##counts.obj  <- countOverlaps(narrow.gr,narrow.gr)
  rjmcmc.df <- rjmcmc.df[!duplicated(rjmcmc.df$start),]
  
  residual.df <- narrow.df[,1:3]
  residual.df <- residual.df[,1:3]
  residual.df$residual <- full.df$residual
  residual.df <- residual.df[!duplicated(residual.df$start),]

  out.list <- list(bed.df,rjmcmc.df,residual.df)
  return(out.list)
}
```

## Promoter files 
```{r}
prom.list <- create_track_dfs(prom.df,prom.full.df)
write.table(x=prom.list[[1]],file=output.dir %&% 
              "promoter_endo_peaky-interactions.bed",sep="\t",row.names=F,
            col.names=F,quote=F)
write.table(x=prom.list[[2]],file=output.dir %&% 
              "promoter_endo_peaky-interactions_rjmcmc-pos.bedGraph",sep="\t",row.names=F,
            col.names=F,quote=F)
write.table(x=prom.list[[3]],file=output.dir %&% 
              "promoter_endo_peaky-interactions_residual.bedGraph",sep="\t",row.names=F,
            col.names=F,quote=F)
```


## Create big bed files with these commands on the command line:
./bedToBigBed bedExample.txt hg38.chrom.sizes myBigBed.bb

```{bash}
sort -k1,1 -k2,2n promoter_endo_peaky-interactions.bed > promoter_endo_peaky-interactions.sorted.bed
./bedToBigBed promoter_endo_peaky-interactions.sorted.bed hg38.chrom.sizes promoter_endo_peaky-interactions.bb
```


## Create bigWig files by converting from the bedGraph files  
./bedGraphToBigWig in.bedGraph chrom.sizes myBigWig.bw
 
NOTE: some of the start/end coordinates had a "+" character in the sorted 
bedGraph files, needed to manually remove these very few lines (none were significant)
```{bash}
sort -k1,1 -k2,2n promoter_endo_peaky-interactions_rjmcmc-pos.bedGraph > promoter_endo_peaky-interactions_rjmcmc-pos.sorted.bedGraph 
./bedGraphToBigWig promoter_endo_peaky-interactions_rjmcmc-pos.sorted.bedGraph hg38.chrom.sizes promoter_endo_peaky-interactions_rjmcmc-pos.bw

sort -k1,1 -k2,2n promoter_endo_peaky-interactions_residual.bedGraph > promoter_endo_peaky-interactions_residual.sorted.bedGraph 
./bedGraphToBigWig promoter_endo_peaky-interactions_residual.sorted.bedGraph hg38.chrom.sizes promoter_endo_peaky-interactions_residual.bw
```

