---
title: "05.2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script will extract values for Table 2 in the manuscript 

Note: HGNC symbol for gene C15orf38-AP3S2 has been replaced with ARPIN-AP3S2
Note: HGNC symbol for gene WARS has been replaced with WARS1

# Setup 

```{r}
"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")
library("GenomicRanges")
rescomp.dir <- "/Users/jasont/science/servers/FUSE5/"
work.dir <- rescomp.dir %&% "projects/wtsa/promoter_driven/"
out.dir <- work.dir %&% "output_files/"
pky.dir <- work.dir %&% "peaky_interactions/"
```

Capture Compare output files 

```{r}
# promoter capture files 
prom.cc.file  <- work.dir %&% "03_CaptureCompare/capture-compare-input_hg38.txt"
prom.cc.dir  <- work.dir %&% "03_CaptureCompare/T2D_Promoter_cis_analysis/"
```


```{r}
cc.df <- fread(prom.cc.file)
cc.df$V1 <- purrr::map(cc.df$V1,function(s){
  gsub("C15orf38-AP3S2","ARPIN-AP3S2",x=s)
}) %>% as.character(.)
cc.df$V1 <- purrr::map(cc.df$V1,function(s){
  gsub("NKX6_3","NKX6-3",x=s)
}) %>% as.character(.)
cc.df$V1 <- purrr::map(cc.df$V1,function(s){
  gsub("WARS","WARS1",x=s)
}) %>% as.character(.)
```

## Lookup peaky chain correlations 

```{r}
corr.df <- fread(pky.dir %&% "promoter-capture_peaky-chain-correlations.txt")
corr.df$capname <- purrr::map(corr.df$capture,function(cap){
  capname <- gsub("C15orf38-AP3S2","ARPIN-AP3S2",cap)
  capname <- gsub("NKX6_3","NKX6-3",capname)
  capname <- gsub("WARS","WARS1",capname)
  return(capname)
}) %>% as.character(.)
res.df <- dplyr::select(corr.df,one_of("capture","capname","correlation")) %>% 
  arrange(capture)
names(res.df)[2] <- "gencodeV37.name"
```

Append genomic region boundaries 

```{r}
genreg.df <- fread(out.dir %&% "genomic-regions_ligation-junctions.txt")
res.df$genomic.region <- purrr::map(res.df$gencodeV37.name,function(g){
  filter(genreg.df,gene==g)$genomic.region
}) %>% as.character(.)
```

## Lookup and append significant peaky interaction information 

```{r}
pky.df <- fread(pky.dir %&% "promoter_endo_peaky-interactions.txt")
```

Subset to results within region (should already be within range, but done as a check)
```{r}
sig.df <- c()
for (cap in unique(pky.df$capture)){
  reg <- filter(res.df,capture==cap)$genomic.region
  chrom <- strsplit(reg,":")[[1]][1]
  vec <- (strsplit(reg,":")[[1]][2] %>% strsplit(.,split="-"))[[1]]
  start.pos <- as.integer(vec[1]); end.pos <- as.integer(vec[2])
  sub.df <- filter(pky.df,capture==cap,seqnames==chrom,start>=start.pos,end<=end.pos)
  sig.df <- rbind(sig.df,sub.df)
}
dim(pky.df)[1]==dim(sig.df)[1] # results are all within the expected range 
```

Append numbers of significant interactions 

```{r}
append_sig_counts <- function(res.df,sig.df){
  prox.count <- c()
  dist.count <- c()
  tot.count <- c()
  pb <- txtProgressBar(min=0,max=dim(res.df)[1],style=3)
  for (i in 1:dim(res.df)[1]){
    setTxtProgressBar(pb,i)
    row.df <- res.df[i,]
    cap <- row.df$capture
    sub.df <- filter(sig.df,capture==cap)
    tot <- dim(sub.df)[1]
    prox <- dim(filter(sub.df,type=="proximal"))[1]
    dist <- dim(filter(sub.df,type=="distal"))[1]
    prox.count <- append(prox.count,prox)
    dist.count <- append(dist.count,dist)
    tot.count <- append(tot.count,tot)    
  }
  out.df <- res.df
  out.df$prox.count <- prox.count
  out.df$dist.count <- dist.count
  out.df$tot.count <- tot.count
  return(out.df)
}
```

```{r}
rescount.df <- append_sig_counts(res.df,sig.df)
all((rescount.df$prox.count + rescount.df$dist.count) == rescount.df$tot.count)
# Sanity check that all interactions are accounted for 
```

## Profile interaction "streches" (i.e. elements) from contiguous sig.frags 

```{r}
get_strech_df <- function(cap,sig.df){
  peak.width.all.vec <- c()
  sub.df <- filter(sig.df,capture==cap)
  sub.gr <- GRanges(seqnames=sub.df$seqnames,
                    IRanges(start=sub.df$start,end=sub.df$end))
  reduce.df <- reduce(sub.gr) %>% as.data.frame()
  reduce.df$peak <- row.names(reduce.df)
  names(reduce.df)[4] <- "peak.width"
  num.peaks <- dim(reduce.df)[1]
  reduce.df$max.mppc <- purrr::map(1:dim(reduce.df)[1],function(i){
    row.df <- reduce.df[i,]
    pk.df <- filter(sub.df,seqnames==row.df$seqnames,
                    start>=row.df$start,end<=row.df$end)
    max(pk.df$rjmcmc_pos)
  }) %>% as.numeric(.)
  peak.width.all.vec <- append(peak.width.all.vec,reduce.df$peak.width)
  med.value <- median(reduce.df$peak.width) %>% round(.,0)
  sd.value <- sd(reduce.df$peak.width) %>% round(.,0)
  med.sd <- med.value %&% "±" %&% sd.value
  med.value2 <- median(reduce.df$max.mppc) %>% round(.,3)
  sd.value2 <- sd(reduce.df$max.mppc) %>% round(.,3)
  med.sd2 <- med.value2 %&% "±" %&% sd.value2
  out.df <- data.frame("capture"=cap,"num.peaks"=num.peaks,
                       "median.width"=med.sd,"median.max.mppc"=med.sd2)
  out.df2 <- reduce.df
  return(list(out.df,out.df2,peak.width.all.vec))
}

build_strech_df <- function(rescount.df,sig.df){
  out.df <- c()
  out.df2 <- c()
  peak.width.all.vec <- c() 
  for (cap in rescount.df$capture){
    build.list <- get_strech_df(cap,sig.df)
    out.df <- rbind(out.df,build.list[[1]])
    out.df2 <- rbind(out.df2,build.list[[2]])
    peak.width.all.vec <- append(peak.width.all.vec,build.list[[3]]) 
  }
  print("Mean peak width across all captures: " %&% mean(peak.width.all.vec))
  print("Median peak width across all captures: " %&% median(peak.width.all.vec))
  print("SD peak width across all captures: " %&% sd(peak.width.all.vec))
  return(list(out.df,out.df2))
}
```

```{r}
resstrecth.list <- build_strech_df(rescount.df,sig.df)
resstrecth.df <- resstrecth.list[[1]]
resstrecth.df <- left_join(rescount.df,resstrecth.df,by="capture")
peaks.df <- resstrecth.list[[2]]

write.table(x=resstrecth.df,file=out.dir %&% "peaky-interaction-summary.txt",
            sep="\t",quote=F,row.names=F,col.names=T)
write.table(x=peaks.df,file=out.dir %&% "peaky-interaction-peaks.txt",
            sep="\t",quote=F,row.names=F,col.names=T)
```


Note: Here are peak length numbers provided by Han Sun for pcHiC experiment 
```{r}
pcvec <- c(10130,17590,17590,12408,6157,8999,8644,22495,14757,9290,
           10573,5075,12456,39049,14117,35350,2669,8305,13838,13168)
```

