---
title: "05.3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script will determine the extent to which peaky interactions at capture 
sites involve trait GWAS SNPs/ eSNPs used in locus selection 


# Setup 

```{r}
"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")
library("GenomicRanges")
rescomp.dir <- "/Users/jasont/science/servers/FUSE5/"
work.dir <- rescomp.dir %&% "projects/wtsa/promoter_driven/"
out.dir <- work.dir %&% "output_files/"
pky.dir <- work.dir %&% "peaky_interactions/"
ref.dir <- work.dir %&% "reference_files/"
```

Read relevant peaky output

```{r}
pky.df <- fread(pky.dir %&% "promoter_endo_peaky-interactions.txt")
peaks.df <- fread(file=out.dir %&% "peaky-interaction-peaks.txt",
            header=T)
summ.df <- fread(file=out.dir %&% "peaky-interaction-summary.txt",
            header=T)
frag.df <- fread(file=out.dir %&% "genomic-regions_frag-info.txt",header=T)
```


## Prepare / process trait variant bed files 

```{r}
eqtl.df <- fread(ref.dir %&% "eQTL-index.bed")
gwas.df <- fread(ref.dir %&% "gwas-index.bed")
rsid.esnp.vec <- purrr::map(eqtl.df$V4,function(s){
  strsplit(s,split="_")[[1]][1]
}) %>% as.character(.)
rsid.gwas.vec <- purrr::map(gwas.df$V4,function(s){
  strsplit(s,split="_")[[1]][1]
}) %>% as.character(.)
```

```{r}
write.table(x=rsid.esnp.vec,file=ref.dir%&%"eSNP-rsids.txt",sep="\t",quote=F,
            row.names=F,col.names=F)
write.table(x=rsid.gwas.vec,file=ref.dir%&%"gwas-rsids.txt",sep="\t",quote=F,
            row.names=F,col.names=F)
```


Submitted these SNP lists to UCSC Table Browser tool (July 5th 2021)
Settings: 
Assembly: December 2013 (GRCh38/hg38)
Group: Variation; Track = All SNPs(151); Table: snp151

Reading in outputted files from browser 
Note: the correct SNP position corresponds to the "chromEnd" column 
```{r}
esnp.ucsc.df <- fread(ref.dir %&% "eSNP-rsid-ucsc-table-browser-query.txt")
gwas.ucsc.df <- fread(ref.dir %&% "gwas-rsid-ucsc-table-browser-query.txt")
```

Updating files to include GRCh38 coordinates 

```{r}
eqtl.df$pos.hg38 <- purrr::map(eqtl.df$V4,function(s){
  rsid <- strsplit(s,split="_")[[1]][1]
  pos <- filter(esnp.ucsc.df,name==rsid)$chromEnd
}) %>% as.integer(.)
eqtl.df$func <- purrr::map(eqtl.df$V4,function(s){
  rsid <- strsplit(s,split="_")[[1]][1]
  pos <- filter(esnp.ucsc.df,name==rsid)$func
}) %>% as.character(.)
gwas.df$pos.hg38 <- purrr::map(gwas.df$V4,function(s){
  rsid <- strsplit(s,split="_")[[1]][1]
  pos <- filter(gwas.ucsc.df,name==rsid)$chromEnd
}) %>% as.integer(.)
gwas.df$func <- NA #purrr::map(gwas.df$V4,function(s){
#  rsid <- strsplit(s,split="_")[[1]][1]
#  pos <- filter(gwas.ucsc.df,name==rsid)$func
#}) %>% as.character(.)
```

Write updated text and bed files with hg38 coordinates 

```{r}
esnp.full.df <- dplyr::select(eqtl.df,one_of("V1","V2","pos.hg38","V4","func"))
gwas.full.df <- dplyr::select(gwas.df,one_of("V1","V2","pos.hg38","V4","func"))
names(esnp.full.df) <- c("chrom","pos.hg19","pos.hg38","name","annotation")
names(gwas.full.df) <- c("chrom","pos.hg19","pos.hg38","name","annotation")
esnp.full.df <- esnp.full.df[!duplicated(esnp.full.df$name),]
gwas.full.df <- gwas.full.df[!duplicated(gwas.full.df$name),]
ensp.bed <- data.frame("V1"=esnp.full.df$chrom,"V2"=esnp.full.df$pos.hg38,
                       "V3"=esnp.full.df$pos.hg38,"V4"=esnp.full.df$name,
                       stringsAsFactors = F)
gwas.bed <- data.frame("V1"=gwas.full.df$chrom,"V2"=gwas.full.df$pos.hg38,
                       "V3"=gwas.full.df$pos.hg38,"V4"=gwas.full.df$name,
                       stringsAsFactors = F)
```

```{r}
write.table(esnp.full.df,file=ref.dir %&% "eQTL-index-hg38.txt",sep="\t",
            quote=F,row.names=F,col.names=T)
write.table(gwas.full.df,file=ref.dir %&% "gwas-index-hg38.txt",sep="\t",
            quote=F,row.names=F,col.names=T)
write.table(ensp.bed,file=ref.dir %&% "eQTL-index-hg38.bed",sep="\t",
            quote=F,row.names=F,col.names=F)
write.table(gwas.bed,file=ref.dir %&% "gwas-index-hg38.bed",sep="\t",
            quote=F,row.names=F,col.names=F)
```


Create big bed files with these commands on the command line:
./bedToBigBed bedExample.txt hg38.chrom.sizes myBigBed.bb

```{bash}
UCSCDIR=/well/mccarthy/users/jason/projects/wtsa/promoter_driven/peaky_interactions/track_files
sort -k1,1 -k2,2n eQTL-index-hg38.bed > eQTL-index-hg38.sorted.bed
$UCSCDIR/bedToBigBed eQTL-index-hg38.sorted.bed $UCSCDIR/hg38.chrom.sizes eQTL-index-hg38.sorted.bb
```

```{bash}
UCSCDIR=/well/mccarthy/users/jason/projects/wtsa/promoter_driven/peaky_interactions/track_files
sort -k1,1 -k2,2n gwas-index-hg38.bed > gwas-index-hg38.sorted.bed
$UCSCDIR/bedToBigBed gwas-index-hg38.sorted.bed $UCSCDIR/hg38.chrom.sizes gwas-index-hg38.sorted.bb
```


## Append position information to colocalisation input file 

```{r}
input.df <- fread(work.dir%&%"input_files/InputForTable3.csv")
temp.df <- c() 
for (i in 1:dim(input.df)[1]){
  row.df <- input.df[i,]
  cap <- row.df$`Captured Gene`
  gwas.snp <- row.df$`GWAS index SNP`
  esnp <- row.df$`Islet eQTL index SNP`
  gwas.pos <- filter(gwas.full.df,grepl(gwas.snp,name))$pos.hg38 %>% unique(.)
  if (cap=="WARS1"){
    gwas.pos<-100365945
  }
  chrom <- filter(esnp.full.df,grepl(esnp,name))$chrom %>% unique(.)
  esnp.pos <- filter(esnp.full.df,grepl(esnp,name))$pos.hg38 %>% unique(.)
  if (cap=="DNLZ"){
    chrom<-"chr9"
    esnp.pos<-136346577 
  }
  build.df <- data.frame("chrom"=chrom,"gwas.pos"=gwas.pos,"esnp.pos"=esnp.pos,
                         stringsAsFactors = F)
  build.df$gwas.pos <- as.integer(build.df$gwas.pos)
  build.df$esnp.pos <- as.integer(build.df$esnp.pos)
  build.df <- cbind(row.df,build.df)
  temp.df <- rbind(temp.df,build.df)
}
input.df <- temp.df

```




## Determine if there is compelling trait interactions at each capture-c locus 


```{r}
extract_info <- function(snp.gr,pky.sub.gr,tss.gr,variant.type){
  # variant.type can be either "esnp" or "gwas" 
  nearest.pair <- nearest(snp.gr,pky.sub.gr)
  nearest.dist <- (distanceToNearest(snp.gr,pky.sub.gr)%>%
    as.data.frame())$distance   
  pky.sub.df <- pky.sub.gr[nearest.pair] %>% as.data.frame()
  nearest.peak.df <- filter(peaks.df,seqnames==pky.sub.df$seqnames,
        start>=pky.sub.df$start,end<=pky.sub.df$end)
  nearest.peak <- (nearest.peak.df$seqnames%&%":"%&%
    nearest.peak.df$start%&%"-"%&%nearest.peak.df$end) %>% unique(.) %>% 
    paste0(.,collapse=",")
  nearest.peak.width <- nearest.peak.df$peak.width %>% unique(.) %>% 
    as.integer(.)
  nearest.peak.max.mppc <- nearest.peak.df$max.mppc %>% unique(.) %>% 
    as.numeric(.)
  nearest.dist.tss <- (distanceToNearest(snp.gr,tss.gr) %>% 
                              as.data.frame())$distance %>% unique(.) %>% 
    as.integer(.)
  name1<-"dist.from."%&%variant.type%&%".to.TSS"
  name2<-"nearest.peak.to."%&%variant.type
  name3<-"dist.from."%&%variant.type%&%".to.nearestPeak"
  name4<-"nearest.peak.to."%&%variant.type%&%".width"
  name5<-"nearest.peak.to."%&%variant.type%&%".maxMPPC"
  out.df <- data.frame(nearest.dist.tss,nearest.peak,nearest.dist,
                nearest.peak.width,nearest.peak.max.mppc,
                stringsAsFactors = FALSE)
  names(out.df) <- c(name1,name2,name3,name4,name5)
  return(out.df)
}
```


```{r}
assess_peaky_overlap <- function(signal.index){
  row.df <- input.df[signal.index,]
  genomic.region <- filter(summ.df,
                        gencodeV37.name==row.df$`Captured Gene`)$genomic.region
  region.coordinates <- (strsplit(x=genomic.region,split=":")[[1]][2] %>% 
    strsplit(.,split="-"))[[1]]
  region.start <- region.coordinates[1] %>% as.integer(.)
  region.end <- region.coordinates[2] %>% as.integer(.)
  pky.sub <- filter(peaks.df,seqnames==row.df$chrom,start>=region.start,
                    end<=region.end)
  pky.sub.gr <- GRanges(seqnames=pky.sub$seqnames,
              IRanges(start=pky.sub$start,end=pky.sub$end))
  frag.sub <- filter(frag.df,gene==row.df$`Captured Gene`)
  tss.vec <- strsplit(frag.sub$tss,split=",")[[1]] %>% as.integer()
  # may have multiple TSS captured by viewpoint 
  tss.gr <- GRanges(seqnames=row.df$chrom,IRanges(start=tss.vec, end=tss.vec)) 
  print(length(pky.sub.gr))
  # eSNP 
  esnp.gr <- GRanges(seqnames=row.df$chrom,
      IRanges(start=row.df$esnp.pos,end=row.df$esnp.pos))
  build.esnp.df <- extract_info(esnp.gr,pky.sub.gr,tss.gr,"esnp")
  # GWAS SNP 
  gwas.gr <- GRanges(seqnames=row.df$chrom,
      IRanges(start=row.df$gwas.pos,end=row.df$gwas.pos))
  build.gwas.df <- extract_info(gwas.gr,pky.sub.gr,tss.gr,"gwas")
  build.df <- cbind(build.esnp.df,build.gwas.df)
  build.df <- cbind(row.df,build.df)
  return(build.df)
}
```


```{r}
build_assessment_df <- function(input.df){
  out.df <- c()
  for (i in 1:dim(input.df)[1]){
    build.df <- assess_peaky_overlap(i)
    out.df <- rbind(out.df,build.df)
  }
  return(out.df)
}
```

```{r}
eval.df <- build_assessment_df(input.df)
esnp.bed.revised <- data.frame("V1"=eval.df$chrom,"V2"=eval.df$esnp.pos,
          "V3"=eval.df$esnp.pos,"V4"=eval.df$`Islet eQTL index SNP`,
          stringsAsFactors = F)
gwas.bed.revised <- data.frame("V1"=eval.df$chrom,"V2"=eval.df$gwas.pos,
          "V3"=eval.df$gwas.pos,"V4"=eval.df$`GWAS index SNP`,
          stringsAsFactors = F)
```


```{r}
write.table(x=eval.df,file=out.dir%&%"integrate-peaky-with-trait-snps.txt",
            sep="\t",quote=F,row.names=F,col.names=T)
write.table(esnp.bed.revised,file=ref.dir %&% "eQTL-index-hg38.revised.bed",
            sep="\t",quote=F,row.names=F,col.names=F)
write.table(gwas.bed.revised,file=ref.dir %&% "gwas-index-hg38.revised.bed",
            sep="\t",quote=F,row.names=F,col.names=F)
```


Create big bed files with these commands on the command line:
./bedToBigBed bedExample.txt hg38.chrom.sizes myBigBed.bb

```{bash}
UCSCDIR=/well/mccarthy/users/jason/projects/wtsa/promoter_driven/peaky_interactions/track_files
sort -k1,1 -k2,2n eQTL-index-hg38.revised.bed > eQTL-index-hg38.revised.sorted.bed
$UCSCDIR/bedToBigBed eQTL-index-hg38.revised.sorted.bed $UCSCDIR/hg38.chrom.sizes eQTL-index-hg38.revised.sorted.bb
```

```{bash}
UCSCDIR=/well/mccarthy/users/jason/projects/wtsa/promoter_driven/peaky_interactions/track_files
sort -k1,1 -k2,2n gwas-index-hg38.revised.bed > gwas-index-hg38.revised.sorted.bed
$UCSCDIR/bedToBigBed gwas-index-hg38.revised.sorted.bed $UCSCDIR/hg38.chrom.sizes gwas-index-hg38.revised.sorted.bb
```




