---
title: "05.3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script will determine the extent to which peaky interactions at capture 
sites involve trait GWAS SNPs/ eSNPs used in locus selection 


# Setup 

```{r}
"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")
library("GenomicRanges")
rescomp.dir <- "/Users/jasont/science/servers/FUSE5/"
work.dir <- rescomp.dir %&% "projects/wtsa/promoter_driven/"
out.dir <- work.dir %&% "output_files/"
pky.dir <- work.dir %&% "peaky_interactions/"
ref.dir <- work.dir %&% "reference_files/"
```

Read relevant peaky output

```{r}
pky.df <- fread(pky.dir %&% "promoter_endo_peaky-interactions.txt")
peaks.df <- fread(file=out.dir %&% "peaky-interaction-peaks.txt",
            header=T)
summ.df <- fread(file=out.dir %&% "peaky-interaction-summary.txt",
            header=T)
frag.df <- fread(file=out.dir %&% "genomic-regions_frag-info.txt",header=T)
```


## Prepare / process trait variant bed files 

```{r}
eqtl.df <- fread(ref.dir %&% "eQTL-index.bed")
gwas.df <- fread(ref.dir %&% "gwas-index.bed")
rsid.esnp.vec <- purrr::map(eqtl.df$V4,function(s){
  strsplit(s,split="_")[[1]][1]
}) %>% as.character(.)
rsid.gwas.vec <- purrr::map(gwas.df$V4,function(s){
  strsplit(s,split="_")[[1]][1]
}) %>% as.character(.)
write.table(x=rsid.esnp.vec,file=ref.dir%&%"eSNP-rsids.txt",sep="\t",quote=F,
            row.names=F,col.names=F)
write.table(x=rsid.gwas.vec,file=ref.dir%&%"gwas-rsids.txt",sep="\t",quote=F,
            row.names=F,col.names=F)
```

Submitted these SNP lists to UCSC Table Browser tool (July 5th 2021)
Settings: 
Assembly: December 2013 (GRCh38/hg38)
Group: Variation; Track = Common SNPs 151; Table: snp151Common

Reading in outputted files from browser 
Note: the correct SNP position corresponds to the "chromEnd" column 
```{r}
esnp.ucsc.df <- fread(ref.dir %&% "eSNP-rsid-ucsc-table-browser-query.txt")
gwas.ucsc.df <- fread(ref.dir %&% "gwas-rsid-ucsc-table-browser-query.txt")
```

Updating files to include GRCh38 coordinates 

```{r}
eqtl.df$pos.hg38 <- purrr::map(eqtl.df$V4,function(s){
  rsid <- strsplit(s,split="_")[[1]][1]
  pos <- filter(esnp.ucsc.df,name==rsid)$chromEnd
}) %>% as.integer(.)
eqtl.df$func <- purrr::map(eqtl.df$V4,function(s){
  rsid <- strsplit(s,split="_")[[1]][1]
  pos <- filter(esnp.ucsc.df,name==rsid)$func
}) %>% as.character(.)
gwas.df$pos.hg38 <- purrr::map(gwas.df$V4,function(s){
  rsid <- strsplit(s,split="_")[[1]][1]
  pos <- filter(gwas.ucsc.df,V4==rsid)$V3
}) %>% as.integer(.)
gwas.df$func <- NA #purrr::map(gwas.df$V4,function(s){
#  rsid <- strsplit(s,split="_")[[1]][1]
#  pos <- filter(gwas.ucsc.df,name==rsid)$func
#}) %>% as.character(.)
```

Write updated text and bed files with hg38 coordinates 

```{r}
esnp.full.df <- dplyr::select(eqtl.df,one_of("V1","V2","pos.hg38","V4","func"))
gwas.full.df <- dplyr::select(gwas.df,one_of("V1","V2","pos.hg38","V4","func"))
names(esnp.full.df) <- c("chrom","pos.hg19","pos.hg38","name","annotation")
names(gwas.full.df) <- c("chrom","pos.hg19","pos.hg38","name","annotation")
esnp.full.df <- esnp.full.df[!duplicated(esnp.full.df$name),]
gwas.full.df <- gwas.full.df[!duplicated(gwas.full.df$name),]
write.table(esnp.full.df,file=ref.dir %&% "eQTL-index-hg38.txt",sep="\t",
            quote=F,row.names=F,col.names=T)
write.table(gwas.full.df,file=ref.dir %&% "gwas-index-hg38.txt",sep="\t",
            quote=F,row.names=F,col.names=T)
ensp.bed <- data.frame("V1"=esnp.full.df$chrom,"V2"=esnp.full.df$pos.hg38,
                       "V3"=esnp.full.df$pos.hg38,"V4"=esnp.full.df$name,
                       stringsAsFactors = F)
gwas.bed <- data.frame("V1"=gwas.full.df$chrom,"V2"=gwas.full.df$pos.hg38,
                       "V3"=gwas.full.df$pos.hg38,"V4"=gwas.full.df$name,
                       stringsAsFactors = F)
write.table(ensp.bed,file=ref.dir %&% "eQTL-index-hg38.bed",sep="\t",
            quote=F,row.names=F,col.names=F)
write.table(gwas.bed,file=ref.dir %&% "gwas-index-hg38.bed",sep="\t",
            quote=F,row.names=F,col.names=F)
```

Create big bed files with these commands on the command line:
./bedToBigBed bedExample.txt hg38.chrom.sizes myBigBed.bb

```{bash}
UCSCDIR=/well/mccarthy/users/jason/projects/wtsa/promoter_driven/peaky_interactions/track_files
sort -k1,1 -k2,2n eQTL-index-hg38.bed > eQTL-index-hg38.sorted.bed
$UCSCDIR/bedToBigBed eQTL-index-hg38.sorted.bed $UCSCDIR/hg38.chrom.sizes eQTL-index-hg38.sorted.bb
```

```{bash}
UCSCDIR=/well/mccarthy/users/jason/projects/wtsa/promoter_driven/peaky_interactions/track_files
sort -k1,1 -k2,2n gwas-index-hg38.bed > gwas-index-hg38.sorted.bed
$UCSCDIR/bedToBigBed gwas-index-hg38.sorted.bed $UCSCDIR/hg38.chrom.sizes gwas-index-hg38.sorted.bb
```



## Determine if there is compelling trait interactions at each capture-c locus 


```{r}
prox5kb.df <- build_5kb_proximity_df()
pky.overlap.df <- build_pky_overlap_df()
```



```{r}
assess_peaky_overlap <- function(cap){
  if (!(cap %in% frag.df$gene)){
    cap <- filter(summ.df,capture==cap)$gencodeV37.name
  }
  lookup.cap <- filter(summ.df,gencodeV37.name==cap)$capture
  lookup.chrom <- (filter(summ.df,gencodeV37.name==cap)$genomic.region %>% 
    strsplit(.,split=":"))[[1]][1]
  lookup.start <- strsplit((filter(summ.df,gencodeV37.name==cap)$genomic.region %>% 
    strsplit(.,split=":"))[[1]][2],split="-")[[1]][1] %>% as.integer(.)  
  lookup.end <- strsplit((filter(summ.df,gencodeV37.name==cap)$genomic.region %>% 
    strsplit(.,split=":"))[[1]][2],split="-")[[1]][2] %>% as.integer(.)
  esnp.sub <- filter(esnp.full.df,chrom==lookup.chrom,pos.hg38>=lookup.start,
                     pos.hg38<=lookup.end)
  gwas.sub <- filter(gwas.full.df,chrom==lookup.chrom,pos.hg38>=lookup.start,
                     pos.hg38<=lookup.end)
  frag.sub <- filter(frag.df,gene==cap)
  tss.chrom <- frag.sub$chrom
  tss.vec <- strsplit(frag.sub$tss,split=",")[[1]] %>% as.integer()
  # may have multiple TSS captured by viewpoint 
  tss.gr <- GRanges(seqnames=tss.chrom,IRanges(start=tss.vec, end=tss.vec))  
  pky.sub <- filter(peaks.df,seqnames==lookup.chrom,start>=lookup.start,
                    end<=lookup.end)
  if (dim(pky.sub)[1]>0 & dim(esnp.sub)[1]>0){
    esnp.gr <- GRanges(seqnames=esnp.sub$chrom,
            IRanges(start=esnp.sub$pos.hg38,end=esnp.sub$pos.hg38))
    names(esnp.gr) <- esnp.sub$name
    pky.sub.gr <- GRanges(seqnames=pky.sub$seqnames,
                IRanges(start=pky.sub$start,end=pky.sub$end))
    nearest.pair <- nearest(esnp.gr,pky.sub.gr)
    nearest.dist <- (distanceToNearest(esnp.gr,pky.sub.gr)%>%as.data.frame())$distance %>%
      unique(.) %>%  paste0(.,collapse=",")
    rsid <- purrr::map(esnp.sub$name,function(s){
        strsplit(s,split="_")[[1]][1]
    }) %>% as.character() 
    egene <- purrr::map(esnp.sub$name,function(s){
      strsplit(s,split="_")[[1]][2]
    }) %>% as.character(.) %>% paste0(.,collapse = ",") 
    snp <- paste0(unique(esnp.sub$chrom%&%":"%&%
           esnp.sub$pos.hg38%&%" ("%&%rsid%&%")"),collapse = ",")
    nearest.peak.df <- pky.sub[nearest.pair]
    nearest.peak.df <- nearest.peak.df[!duplicated(nearest.peak.df$start),]
    nearest.peak <- (nearest.peak.df$seqnames%&%":"%&%
      nearest.peak.df$start%&%"-"%&%nearest.peak.df$end) %>% unique(.) %>% 
      paste0(.,collapse=",")
    nearest.peak.width <- nearest.peak.df$peak.width %>% unique(.) %>% 
      paste0(.,collapse=",")
    nearest.peak.max.mppc <- nearest.peak.df$max.mppc %>% unique(.) %>% 
      paste0(.,collapse=",")
    nearest.dist.esnp.tss <- (distanceToNearest(esnp.gr,tss.gr) %>% 
                                as.data.frame())$distance %>% unique(.) %>% 
      paste0(.,collapse=",")
    nearest.dist.esnp.tss <- (distanceToNearest(esnp.gr,tss.gr) %>% 
                                as.data.frame())$distance %>% unique(.) %>% 
      paste0(.,collapse=",")
    esnp.build <- data.frame("capture"=cap,"eSNP"=snp,"eGene"=egene,
                        "dist.from.eSNP.to.TSS"=nearest.dist.esnp.tss,
                        "nearest.peak.to.eSNP"=nearest.peak,
                        "dist.from.eSNP.to.nearestPeak"=nearest.dist,
                        "nearest.peak.to.eSNP.width"=nearest.peak.width,
                        "nearest.peak.to.eSNP.maxMPPC"=nearest.peak.max.mppc,
                        stringsAsFactors = FALSE)
  } else{
    esnp.build <- data.frame("capture"=cap,"eSNP"=NA,"eGene"=NA,
                    "dist.from.eSNP.to.TSS"=NA,
                    "nearest.peak.to.eSNP"=NA,
                    "dist.from.eSNP.to.nearestPeak"=NA,
                    "nearest.peak.to.eSNP.width"=NA,
                    "nearest.peak.to.eSNP.maxMPPC"=NA,
                    stringsAsFactors = FALSE)
  }
  if (dim(pky.sub)[1]>0 & dim(gwas.sub)[1]>0){
    gwas.gr <- GRanges(seqnames=gwas.sub$chrom,
            IRanges(start=gwas.sub$pos.hg38,end=gwas.sub$pos.hg38))
    names(gwas.gr) <- gwas.sub$name
    pky.sub.gr <- GRanges(seqnames=pky.sub$seqnames,
                IRanges(start=pky.sub$start,end=pky.sub$end))
    nearest.pair <- nearest(gwas.gr,pky.sub.gr)
    nearest.dist <- (distanceToNearest(gwas.gr,pky.sub.gr)%>%as.data.frame())$distance %>% 
      unique(.) %>%  paste0(.,collapse=",")
    rsid <- purrr::map(gwas.sub$name,function(s){
      strsplit(s,split="_")[[1]][1]
    }) %>% as.character() 
    trait <- purrr::map(gwas.sub$name,function(s){
      svec <- strsplit(s,split="_")[[1]]
      paste0(svec[2:length(svec)],collapse="-")
    }) %>% as.character(.) %>% paste0(.,collapse = ",") 
    snp <- paste0(unique(gwas.sub$chrom%&%":"%&%
           gwas.sub$pos.hg38%&%" ("%&%rsid%&%")"),collapse = ",")
    nearest.peak.df <- pky.sub[nearest.pair]
    nearest.peak <- (nearest.peak.df$seqnames%&%":"%&%
      nearest.peak.df$start%&%"-"%&%nearest.peak.df$end) %>% unique(.) %>% 
      paste0(.,collapse=",")
    nearest.peak.width <- nearest.peak.df$peak.width %>% unique(.) %>% 
      paste0(.,collapse=",")
    nearest.peak.max.mppc <- nearest.peak.df$max.mppc %>% unique(.) %>% 
      paste0(.,collapse=",")
    nearest.dist.gwas.tss <- (distanceToNearest(gwas.gr,tss.gr) %>% 
                                as.data.frame())$distance %>% unique(.) %>% 
      paste0(.,collapse=",")
    gwas.build <- data.frame("capture"=cap,"gwas"=snp,"trait"=trait,
                        "dist.from.gwas.to.TSS"=nearest.dist.gwas.tss,
                        "nearest.peak.to.gwas"=nearest.peak,
                        "dist.from.gwas.to.nearestPeak"=nearest.dist,
                        "nearest.peak.to.gwas.width"=nearest.peak.width,
                        "nearest.peak.to.gwas.maxMPPC"=nearest.peak.max.mppc,
                        stringsAsFactors = FALSE)
  } else{
    gwas.build <- data.frame("capture"=cap,"gwas"=NA,"trait"=NA,
                    "dist.from.gwas.to.TSS"=NA,
                    "nearest.peak.to.gwas"=NA,
                    "dist.from.gwas.to.nearestPeak"=NA,
                    "nearest.peak.to.gwas.width"=NA,
                    "nearest.peak.to.gwas.maxMPPC"=NA,
                    stringsAsFactors = FALSE)
  }
  out.df <- dplyr::inner_join(esnp.build,gwas.build,by="capture")
  return(out.df)
}

build_assessment_df <- function(){
  out.df <- c()
  for (cap in summ.df$capture){
    print(cap)
    build.df <- assess_peaky_overlap(cap)
    out.df <- rbind(out.df,build.df)
  }
  return(out.df)
}
```

```{r}
eval.df <- build_assessment_df()
write.table(x=eval.df,file=out.dir%&%"integrate-peaky-with-trait-snps.txt",
            sep="\t",quote=F,row.names=F,col.names=T)
```




