---
title: '06.1'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup 

```{r}
"%&%" <- function(a,b) paste0(a,b)
library("data.table");library("tidyverse");library("viridis")
serv.dir <- "/Users/jasont/science/"
work.dir <- serv.dir %&% "projects/wtsa/promoter_driven/"
table.dir <- work.dir %&% "result_tables/"
figure.dir <- table.dir %&% "figures/"
tab.file <- table.dir %&% "Tables-for-promoter-capture-manuscript_Table1.csv"
tab.df <- fread(tab.file,header=T,skip=2) %>% as.data.frame()
```

## Evaluate read counts per capture across loci 

Create and process ligation junction dataframe for comparisons 
```{r}
junction.df <- cbind(tab.df[,1],
  tab.df[,grepl("Number of mapped reads",x=names(tab.df))])
names(junction.df) <- c("Gene/Locus","V2","V3","counts.hic.endo",
                     "counts.hic.islet")
junction.df$V3 <- purrr::map(junction.df$V3,function(s){
  gsub(x=s,"+-","±",fixed = TRUE)
}) %>% as.character(.)
junction.df$counts.capc.endo <- purrr::map(junction.df$V2,function(s){
  strsplit(s,split="±")[[1]][1]
}) %>% as.integer(.)
junction.df$sd.capc.endo <- purrr::map(junction.df$V2,function(s){
  strsplit(s,split="±")[[1]][2]
}) %>% as.integer(.)
junction.df$counts.pchic.islet <- purrr::map(junction.df$V3,function(s){
  strsplit(s,split="±")[[1]][1]
}) %>% as.integer(.)
junction.df$sd.pchic.islet <- purrr::map(junction.df$V3,function(s){
  strsplit(s,split="±")[[1]][2]
}) %>% as.integer(.)
```

```{r}
write.table(x=junction.df,file=work.dir %&% "result_tables/"%&%
              "junction-counts.txt",row.names=F,col.names=T,
            sep="\t",quote=F)
```


Reformat ligation junction dataframe for barplot 
```{r}
plot.df1 <- c()
for (i in 1:dim(junction.df)[1]){
  sub.df <- junction.df[i,]
  build.df <- data.frame("Gene.Locus"=sub.df$`Gene/Locus`,
  "Count"=c(sub.df$counts.capc.endo,sub.df$counts.pchic.islet,
            sub.df$counts.hic.endo,sub.df$counts.hic.islet),
  "sd"=c(sub.df$sd.capc.endo,sub.df$sd.pchic.islet,0,0),
  "Experiment"=c("NG-CapC: endoC","PC-HiC: islet",
                 "HiC: endoC","HiC; islet"))
  plot.df1 <- rbind(plot.df1,build.df)
}
plot.df1$Experiment <- factor(plot.df1$Experiment,
  levels=c("HiC: endoC","HiC; islet",
           "PC-HiC: islet","NG-CapC: endoC"))
plot.df1$Count <- plot.df1$Count/1000
plot.df1$sd <- plot.df1$sd/1000
```

Create ligation junction barplot, with comparison across captures 
```{r}
plt1 <- ggplot(data=plot.df1,
  aes_string(x="Gene.Locus",y="Count",fill="Experiment")) + 
  geom_hline(yintercept = c(10,20,30),color="gray50",linetype=2) + 
  geom_bar(stat="identity",position=position_dodge(),
           color="black",size=0.5) +
  ##geom_errorbar(aes(x=Gene.Locus,ymin=Count-sd,ymax=Count+sd,color=Experiment),
  ##              stat="identity",position=position_dodge(),
  ##         size=0.5) + # color="black",
  scale_fill_manual(values=viridis(10)[c(2,4,6,9)]) +
  ###scale_color_manual(values=viridis(10)[c(2,4,6,9)]) + 
  scale_y_continuous(breaks=seq(0,42,2)) + 
  theme(axis.text.x=element_text(angle=45,vjust=0.6),
        panel.border = element_blank(),
        panel.grid.major.y = element_blank(),#line(color="gray90"),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black",size = rel(1)),
        legend.key = element_blank(),
        panel.background = element_rect(fill="white",color="white"),
        strip.background = element_rect(fill = "white", colour = "black", 
                size = rel(2)), complete = TRUE) + 
  ylab("Number of ligation junctions (thousands)") + 
  xlab("Captured Gene") + 
  ggtitle("Mapped ligation junctions across experiments") + 
  #theme_classic() + 
  coord_cartesian(expand = FALSE)
```

```{r}
ggsave(plot=plt1,filename=figure.dir%&%"junctions-captures.svg",
       width=14,height=4)
ggsave(plot=plt1,filename=figure.dir%&%"junctions-captures.png",
       width=14,height=4)
```

Compare the average counts across experiments 
```{r}
junction.mean.df <- data.frame("Experiment"=c("NG-CapC: endoC","PC-HiC: islet",
                 "HiC: endoC","HiC; islet"),
                 "mean.counts"=c(mean(junction.df$counts.capc.endo),
                                 mean(junction.df$counts.pchic.islet),
                                 mean(junction.df$counts.hic.endo),
                                 mean(junction.df$counts.hic.islet)),
                 "sd.counts"=c(mean(junction.df$sd.capc.endo),
                                 mean(junction.df$sd.pchic.islet),
                                 0,0))
junction.mean.df$Experiment <- factor(junction.mean.df$Experiment,
  levels=c("HiC: endoC","HiC; islet",
           "PC-HiC: islet","NG-CapC: endoC"))

```

Create ligation junction barplot, with mean values
```{r}
plt2 <- ggplot(data=junction.mean.df,
  aes_string(x="Experiment",y="mean.counts")) + 
  geom_errorbar(aes(ymin=(mean.counts-sd.counts),ymax=mean.counts+sd.counts),
                size=0.5,width=0.1) +
  geom_point(shape=21,fill=viridis(10)[4],color="black",size=2) + 
  #scale_fill_manual(values=viridis(10)[c(2,4,6,9)]) + 
  scale_y_continuous(breaks=c(1000,seq(0,30000,5000))) + 
  theme(axis.text.x=element_text(angle=45,vjust=0.9,hjust=1),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black",size = rel(1)),
        legend.key = element_blank(),
        panel.background = element_rect(fill="white",color="white"),
        strip.background = element_rect(fill = "white", colour = "black", 
                size = rel(2)), complete = TRUE) + 
  ylab("Mean number of ligation junctions") + 
  xlab("Experiment") + 
  ggtitle("Mapped ligation junctions across experiments") + 
  #theme_classic() + 
  coord_cartesian(expand = FALSE)
```

```{r}
ggsave(plot=plt2,filename=figure.dir%&%"junctions-captures-mean-sd.pdf",
       width=5,height=4)
ggsave(plot=plt2,filename=figure.dir%&%"junctions-captures-mean-sd.png",
       width=5,height=4)
```

```{r}
plot.df1$Count <- plot.df1$Count * 1000
```

Quantify differences in mean counts of ligation junctions across experiments
```{r}
hic.islet <- junction.mean.df$mean.counts[4]
hic.endo <- junction.mean.df$mean.counts[3]
pchic.islet <- junction.mean.df$mean.counts[2]
capc.islet <- junction.mean.df$mean.counts[1]

ratio1 <- hic.endo/hic.islet # 1.9x more hiC reads in endos vs islets 
ratio2a <- pchic.islet/hic.endo # 2.5x more reads in pcHiC islets vs hiC endos 
ratio2b <- pchic.islet/hic.islet # 4.8x more reads in pcHiC islets vs hiC islets 
ratio3a <- capc.islet/hic.endo # 38.3x more reads in NG-CapC endos vs hiC endos
ratio3b <- capc.islet/hic.islet # 73.3x more reads in NG-CapC endos vs hiC islets
ratio4 <- capc.islet/pchic.islet # 15.3 more reads in NG-CapC endos vs pcHiC islets
```

## Assess restriction fragment count distribution 

```{r}
frag.count.df <- cbind(tab.df[,1],
  tab.df[,grepl("Number of restriction fragments",x=names(tab.df))])
frag.count.plot.df <- c()
for (i in 1:dim(frag.count.df)[1]){
  sub.df <- frag.count.df[i,]
  build.df <- data.frame("Gene"=sub.df$`tab.df[, 1]`,
  "Count"=c(sub.df$`Number of restriction fragments (DpnII)`,
            sub.df$`Number of restriction fragments (HindIII)`),
  "RE"=c("DpnII","HindIII"))
  frag.count.plot.df <- rbind(frag.count.plot.df,build.df)
}
```

```{r}
plt3 <- ggplot(data=frag.count.plot.df,aes(x=Count,fill=RE)) + 
  geom_histogram(color="black",binwidth=100) + 
  scale_fill_manual(values=viridis(10)[c(3,6)]) + 
  scale_x_continuous(breaks=seq(0,7000,500)) + 
  scale_y_continuous(breaks=seq(0,8,1)) + 
  ylab("Count") + xlab("Number of restriction fragments") + 
  theme_classic()
```

```{r}
ggsave(plot=plt3,filename=figure.dir%&%
         "number-fragments-hist.pdf",
       width=6,height=3)
ggsave(plot=plt3,filename=figure.dir%&%
         "number-fragments-hist.png",
       width=6,height=3)
```

```{r}
mean.dpnII<-frag.count.df$`Number of restriction fragments (DpnII)` %>% 
  mean(.) # 5248.111
mean.hindIII<-frag.count.df$`Number of restriction fragments (HindIII)` %>% 
  mean(.) # 514.1481
ratio.frag <- mean.dpnII/mean.hindIII # 10.21x more fragments
```

Lookup mean of mean fragment lenghts (DpnII vs HindIII)
```{r}
mean.frag.len.df <- cbind(tab.df[,1],
  tab.df[,grepl("Mean fragment length",x=names(tab.df))])[,1:3]
names(mean.frag.len.df) <- c("Gene","mean.frag.len.dpnII",
                             "mean.frag.len.hindIII")
mean.frag.len.df$mean.frag.len.hindIII <- purrr::map(mean.frag.len.df$mean.frag.len.hindIII,
  function(s){gsub(x=s,"+-","±",fixed = TRUE)
}) %>% as.character(.)
mean.frag.len.df$mean.dpnII <- purrr::map(mean.frag.len.df$mean.frag.len.dpnII,function(s){
  strsplit(s,split="±")[[1]][1]
}) %>% as.integer(.)
mean.frag.len.df$sd.dpnII <- purrr::map(mean.frag.len.df$mean.frag.len.dpnII,function(s){
  strsplit(s,split="±")[[1]][2]
}) %>% as.integer(.)
mean.frag.len.df$mean.hindIII <- purrr::map(mean.frag.len.df$mean.frag.len.hindIII,function(s){
  strsplit(s,split="±")[[1]][1]
}) %>% as.integer(.)
mean.frag.len.df$sd.hindIII <- purrr::map(mean.frag.len.df$mean.frag.len.hindIII,function(s){
  strsplit(s,split="±")[[1]][2]
}) %>% as.integer(.)
```

```{r}
mean(mean.frag.len.df$mean.dpnII) # Mean fragment length: DpnII - 385.44
mean(mean.frag.len.df$mean.hindIII) # Mean fragment length: hindIII- 4107.30

```




### Mapped reads per restriction fragment 

```{r}
junct.per.frag.df <- cbind(tab.df[,1],
  tab.df[,grepl("Mapped reads per restriction fragment",
                x=names(tab.df))])
names(junct.per.frag.df) <- c("V1","V2","V3","V4","V5")
junct.per.frag.df$V3 <- purrr::map(junct.per.frag.df$V3,function(s){
  gsub(x=s,"+-","±",fixed = TRUE)
}) %>% as.character(.)
junct.per.frag.df$V4 <- purrr::map(junct.per.frag.df$V4,function(s){
  gsub(x=s,"+-","±",fixed = TRUE)
}) %>% as.character(.)
junct.per.frag.df$V5 <- purrr::map(junct.per.frag.df$V5,function(s){
  gsub(x=s,"+-","±",fixed = TRUE)
}) %>% as.character(.)

junct.per.frag.df$counts.capc.endo <- purrr::map(junct.per.frag.df$V2,function(s){
  strsplit(s,split="±")[[1]][1]
}) %>% as.numeric(.)
junct.per.frag.df$sd.capc.endo <- purrr::map(junct.per.frag.df$V2,function(s){
  strsplit(s,split="±")[[1]][2]
}) %>% as.numeric(.)

junct.per.frag.df$counts.pchic.islet <- purrr::map(junct.per.frag.df$V3,function(s){
  strsplit(s,split="±")[[1]][1]
}) %>% as.numeric(.)
junct.per.frag.df$sd.pchic.islet <- purrr::map(junct.per.frag.df$V3,function(s){
  strsplit(s,split="±")[[1]][2]
}) %>% as.numeric(.)

junct.per.frag.df$counts.hic.endo <- purrr::map(junct.per.frag.df$V4,function(s){
  strsplit(s,split="±")[[1]][1]
}) %>% as.numeric(.)
junct.per.frag.df$sd.hic.endo <- purrr::map(junct.per.frag.df$V4,function(s){
  strsplit(s,split="±")[[1]][2]
}) %>% as.numeric(.)

junct.per.frag.df$counts.hic.islet <- purrr::map(junct.per.frag.df$V5,function(s){
  strsplit(s,split="±")[[1]][1]
}) %>% as.numeric(.)
junct.per.frag.df$sd.hic.islet <- purrr::map(junct.per.frag.df$V5,function(s){
  strsplit(s,split="±")[[1]][2]
}) %>% as.numeric(.)
```

Reformat ligation junctions per fragment dataframe for barplot 
```{r}
plot.df2 <- c()
for (i in 1:dim(junct.per.frag.df)[1]){
  sub.df <- junct.per.frag.df[i,]
  build.df <- data.frame("Gene.Locus"=sub.df$V1,
  "Count"=c(sub.df$counts.capc.endo,sub.df$counts.pchic.islet,
            sub.df$counts.hic.endo,sub.df$counts.hic.islet),
  "sd"=c(sub.df$sd.capc.endo,sub.df$sd.pchic.islet,0,0),
  "Experiment"=c("NG-CapC: endoC","PC-HiC: islet",
                 "HiC: endoC","HiC; islet"))
  plot.df2 <- rbind(plot.df2,build.df)
}
plot.df2$Experiment <- factor(plot.df2$Experiment,
  levels=c("HiC: endoC","HiC; islet",
           "PC-HiC: islet","NG-CapC: endoC"))
```

Create ligation junction barplot, with comparison across captures 
```{r}
plt4 <- ggplot(data=plot.df2,
  aes_string(x="Gene.Locus",y="Count",fill="Experiment")) + 
  geom_bar(stat="identity",position=position_dodge(),
           color="black",size=0.5) +
  scale_fill_manual(values=viridis(10)[c(2,4,6,9)]) + 
  #scale_y_continuous(breaks=seq(0,42,2)) + 
  theme(axis.text.x=element_text(angle=45,vjust=0.6),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black",size = rel(1)),
        legend.key = element_blank(),
        panel.background = element_rect(fill="white",color="white"),
        strip.background = element_rect(fill = "white", colour = "black", 
                size = rel(2)), complete = TRUE) + 
  ylab("Mean ligation junctions per fragment") + 
  xlab("Captured Gene") + 
  ggtitle("Ligation junctions per fragment across experiments") + 
  #theme_classic() + 
  coord_cartesian(expand = FALSE)
```

```{r}
ggsave(plot=plt4,filename=figure.dir%&%"junctions-per-frag.svg",
       width=6,height=5)
ggsave(plot=plt4,filename=figure.dir%&%"junctions-per-frag.png",
       width=6,height=5)
```


Compare the average counts across experiments 
```{r}
jpf.mean.df <- data.frame("Experiment"=c("NG-CapC: endoC","PC-HiC: islet",
                 "HiC: endoC","HiC; islet"),
                 "mean.counts"=c(mean(junct.per.frag.df$counts.capc.endo),
                                 mean(junct.per.frag.df$counts.pchic.islet),
                                 mean(junct.per.frag.df$counts.hic.endo),
                                 mean(junct.per.frag.df$counts.hic.islet)),
                 "sd.counts"=c(mean(junct.per.frag.df$sd.capc.endo),
                                 mean(junct.per.frag.df$sd.pchic.islet),
                                 0,0))
jpf.mean.df$Experiment <- factor(jpf.mean.df$Experiment,
  levels=c("HiC: endoC","HiC; islet",
           "PC-HiC: islet","NG-CapC: endoC"))

```

Create ligation junction barplot, with mean values
```{r}
plt5 <- ggplot(data=jpf.mean.df,
  aes_string(x="Experiment",y="mean.counts")) + 
  geom_errorbar(aes(ymin=(mean.counts-sd.counts),ymax=mean.counts+sd.counts),
                size=0.5,width=0.1) +
  geom_point(shape=21,fill=viridis(10)[4],color="black",size=2) + 
  #scale_fill_manual(values=viridis(10)[c(2,4,6,9)]) + 
  #scale_y_continuous(breaks=c(1000,seq(0,30000,5000))) + 
  theme(axis.text.x=element_text(angle=45,vjust=0.9,hjust=1),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black",size = rel(1)),
        legend.key = element_blank(),
        panel.background = element_rect(fill="white",color="white"),
        strip.background = element_rect(fill = "white", colour = "black", 
                size = rel(2)), complete = TRUE) + 
  ylab("Mean number of ligation junctions per fragment") + 
  xlab("Experiment") + 
  ggtitle("Junctions per fragment across experiments") + 
  #theme_classic() + 
  coord_cartesian(expand = FALSE)
```

```{r}
ggsave(plot=plt5,filename=figure.dir%&%"junctions-per-frag-mean-sd.svg",
       width=5,height=4)
ggsave(plot=plt5,filename=figure.dir%&%"junctions-per-frag-mean-sd.png",
       width=5,height=4)
```

