---
title: "06.3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup 

```{r}
"%&%" <- function(a,b) paste0(a,b)
library("data.table");library("tidyverse");library("viridis")
serv.dir <- "/Users/jasont/science/"
work.dir <- serv.dir %&% "projects/wtsa/promoter_driven/"
table.dir <- work.dir %&% "result_tables/"
figure.dir <- table.dir %&% "figures/"
tab.file <- table.dir %&% "Tables-for-promoter-capture-manuscript_Table3.csv"
tab.df <- fread(tab.file,header=T,skip=2) %>% as.data.frame()
tab.df <- tab.df[,1:17]
```

11,15

Summary concordance of interactions and genetic signals 
```{r}
tab.df[,11] %>% table(.)
tab.df[,15] %>% table(.)
```
Of the 25 eQTL/GWAS variant signals, 10 are supported by capC interactions and 7 are supported by pcHiC interactions (w/n 500 bps) 
Note: there were failed peaky convergences in the pcHiC experiment at genes DGKB,ERAP2,GCK,GPSM1,LARP6,TNFSF11
Moreover, NKX6-3 did converge, but no interactions were significant. This accounts for the NAs observed for the pcHiC experiment 

Lookup the concordant signals 
```{r}
names(tab.df)[c(11,15)] <- c("sig.capc.endo",
                             "sig.pchic.islet")
tab.df$sig.pchic.islet[is.na(tab.df$sig.pchic.islet)] <- "NA"
both.df <- filter(tab.df,sig.capc.endo=="Yes",
                       sig.pchic.islet=="Yes")#5
capc.only.df <- filter(tab.df,sig.capc.endo=="Yes",
                       sig.pchic.islet!="Yes")# 5 
pchic.only.df <- filter(tab.df,sig.pchic.islet=="Yes",
                       sig.capc.endo!="Yes")# 2 
```

Note: There are five genetic signals supported by both ng-capC (endos) and 
pcHiC (islets): 
          Locus Captured Gene GWAS SNP rsid GWAS SNP position (GRCh38)  eSNP rsid
1 CDC123/CAMK1D        CAMK1D    rs11257655             chr10:12265895 rs11257658
2         GPSM1         CARD9    rs28505901             chr9:136346577 rs61386106
3        KCNK16        KCNK17     rs1535500              chr6:39316274  rs3807042
4          MADD          MADD    rs10501320             chr11:47272248 rs11039165
5         WARS1         WARS1     rs2146105            chr14:100365945  rs2146105
  eSNP position (GRCh38)
1         chr10:12267269
2         chr9:136352316
3          chr6:39323058
4         chr11:47291138
5        chr14:100365945

Note: There are five genetic signals supported only in ng-capC (endos)
  Locus Captured Gene GWAS SNP rsid GWAS SNP position (GRCh38)  eSNP rsid
1 ADCY5         ADCY5    rs11708067             chr3:123346931 rs11708067
2  DGKB          DGKB    rs17168486              chr7:14858657 rs17168486
3 GPSM1         GPSM1    rs28505901             chr9:136346577 rs28642213
4 GPSM1         GPSM1    rs78403475             chr9:136341149 rs28642213
5 GPSM1          DNLZ    rs28505901             chr9:136346577 rs28505901
  eSNP position (GRCh38)
1         chr3:123346931
2          chr7:14858657
3         chr9:136353630
4         chr9:136353630
5         chr9:136346577

Note: There are two genetic signals supported only in pcHiC (islets)
   Locus Captured Gene GWAS SNP rsid GWAS SNP position (GRCh38)  eSNP rsid
1 TCF7L2        TCF7L2     rs7903146            chr10:112998590  rs7903146
2 UBE2E2        UBE2E2    rs35352848              chr3:23414091 rs13094957
  eSNP position (GRCh38)
1        chr10:112998590
2          chr3:23415589