---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup 


```{r}

"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")


local.dir <- "/home/jason/science/projects/wtsa/joint_analyses/"
rescomp.dir <- "/home/jason/science/servers/FUSE5/"
cbrg.dir <- "/home/jason/science/servers/FUSE2/"
work.dir <- rescomp.dir %&% "projects/wtsa/joint_analyses/"
file.dir <- work.dir %&% "analysis_files/"
loc.df <- fread(file.dir %&% "locus_captures.txt")
full.df <- fread(file.dir %&% "all_locus_interactions_merge1K.txt")
```


```{r}

library("rtracklayer")
gtf <- rtracklayer::import(rescomp.dir %&% "datasets/gencode.v19.genes.v7.patched_contigs.gtf")
gtf.df=as.data.frame(gtf)
pb <- txtProgressBar(min=0,max=dim(gtf.df)[1],style=3)
gtf.df$ensgene <- map(1:dim(gtf.df)[1],function(i){
  setTxtProgressBar(pb,i)
  ens <- gtf.df$gene_id[i]
  (strsplit(ens,split=".",fixed=TRUE))[[1]][1]
}) %>% as.character(.)

```



```{r}


loc.df$locus[!(loc.df$locus %in% unique(full.df$locus))]
full.df$method %>% table(.)
full.df$type %>% table(.)
(filter(full.df,method=="peakC",peakc.deseq2.padj<0.05) %>% dim(.))[1] # 546
all.genes <- map(full.df$genes,function(g){
  strsplit(g,split=",")
}) %>% unlist(.) %>% unique(.)

islet.tpm.df <- fread(rescomp.dir %&% "projects/t2d_classification/analysis_files/islet_expression_tpm.txt")
islet.ess.df <- fread(rescomp.dir %&% "projects/t2d_classification/analysis_files/expression_specificity_scores.txt")
islet.tpm.df$GeneID <- map(1:dim(islet.tpm.df)[1],function(i){
  ens <- islet.tpm.df$GeneID[i]
  (strsplit(ens,split=".",fixed=TRUE))[[1]][1]
}) %>% as.character(.)
islet.ess.df$GeneID <- map(1:dim(islet.ess.df)[1],function(i){
  ens <- islet.ess.df$GeneID[i]
  (strsplit(ens,split=".",fixed=TRUE))[[1]][1]
}) %>% as.character(.)


pb <- txtProgressBar(min=0,max=length(all.genes),style=3)
gene.name <- map(1:length(all.genes),function(i){
  g <- all.genes[i]
  setTxtProgressBar(pb,i)
  filter(gtf.df,ensgene==g)$gene_name %>% unique(.) 
}) %>% as.character(.)

pb <- txtProgressBar(min=0,max=length(all.genes),style=3)
gene.type <- map(1:length(all.genes),function(i){
  g <- all.genes[i]
  setTxtProgressBar(pb,i)
  filter(gtf.df,ensgene==g)$gene_type %>% unique(.) %>% paste0(.,collapse=",")
}) %>% as.character(.)

gene.df <- data.frame("gene"=all.genes,"gene.name"=gene.name,"gene.type"=gene.type,stringsAsFactors = F)
gene.df$gene.type %>% table(.) %>% sort(.) 

mean.islet.tpm <- map(1:length(all.genes),function(i){
  g <- all.genes[i]
  setTxtProgressBar(pb,i)
  filter(islet.tpm.df,GeneID==g)[1,3:dim(islet.tpm.df)[2]] %>% as.numeric(.) %>% mean(.) # will need to redo mapping to ENSID, as Y_RNA is too non-specific 
}) %>% as.numeric(.)
median.islet.tpm <- map(1:length(all.genes),function(i){
  g <- all.genes[i]
  setTxtProgressBar(pb,i)
  filter(islet.tpm.df,GeneID==g)[1,3:dim(islet.tpm.df)[2]] %>% as.numeric(.) %>% median(.) # will need to redo mapping to ENSID, as Y_RNA is too non-specific 
}) %>% as.numeric(.)
islet.ess <- map(1:length(all.genes),function(i){
  g <- all.genes[i]
  setTxtProgressBar(pb,i)
  (filter(islet.ess.df,GeneID==g) %>% arrange(.,desc(islet.score)))[1,]$islet.score
}) %>% as.numeric(.)

gene.df <- data.frame("gene"=all.genes,"gene.name"=gene.name,"gene.type"=gene.type,
                      mean.islet.tpm,median.islet.tpm,islet.ess,stringsAsFactors = F)
gene.df$gene.type %>% table(.) %>% sort(.) 
gene.df$islet.ess %>% na.omit(.) %>% summary(.)
filter(gene.df,gene.type=="protein_coding")$islet.ess %>% na.omit(.) %>% summary(.)

```


```{r}

gene.df$method <- map(1:dim(gene.df)[1],function(i){
  ens <- gene.df$gene[i]
  sub <- filter(full.df,grepl(x=genes,pattern=ens))
  vec <- sub$method %>% unique(.)
  ifelse(length(vec)==2,"both",
         ifelse(vec=="DESeq2","DESeq2",
                ifelse(any(na.omit(sub$peakc.deseq2.padj)<0.05),"peakC.deseq2","peakC")))
                   
}) %>% as.character(.)

write.table(x=gene.df,file=file.dir%&%"all_interaction_genes.txt",sep="\t",row.names=F,quote=F)


```


ESS enrichment 

```{r}

gene_set_enrich <- function(query.vec,full.vec,iter=10000){
  obs <- filter(islet.ess.df,GeneID%in%query.vec)$islet.score %>% na.omit(.) %>% mean(.)
  null.vec <- c()
  pb <- txtProgressBar(min=0,max=iter,style=3)
  for (i in 1:iter){
    setTxtProgressBar(pb,i)
    null.set <- sample(full.vec,size=length(query.vec),replace=FALSE)
    null.count <- filter(islet.ess.df,GeneID%in%null.set)$islet.score %>% na.omit(.) %>% mean(.)
    null.vec <- append(null.vec,null.count)
  }
  enrich.factor <- obs/mean(null.vec)
  emp.p <- (sum(null.vec >= obs) + 1) / (iter+1)
  return(list(enrich.factor,emp.p))
}


```


```{r}

set.seed(1)
test1 <- gene_set_enrich(query.vec=(filter(gene.df,gene.type=="protein_coding")$gene %>% unique(.)),
                         full.vec=(filter(gtf.df,gene_type=="protein_coding")$ensgene %>% unique(.))) # 275 genes; EF = 0.988; pval=0.60
test2 <- gene_set_enrich(query.vec=(filter(gene.df,gene.type=="protein_coding",method=="peakC")$gene %>% unique(.)),
                         full.vec=(filter(gtf.df,gene_type=="protein_coding")$ensgene %>% unique(.))) # 65 genes; EF = 0.97; pval=0.61
test3 <- gene_set_enrich(query.vec=(filter(gene.df,gene.type=="protein_coding",method=="DESeq2")$gene %>% unique(.)),
                         full.vec=(filter(gtf.df,gene_type=="protein_coding")$ensgene %>% unique(.))) # 76 genes; EF = 1.2; pval = 0.023
test4 <- gene_set_enrich(query.vec=(filter(gene.df,gene.type=="protein_coding",method=="both")$gene %>% unique(.)),
                         full.vec=(filter(gtf.df,gene_type=="protein_coding")$ensgene %>% unique(.))) # 58 genes; EF = 0.93; pval = 0.76
test5 <- gene_set_enrich(query.vec=(filter(gene.df,gene.type=="protein_coding",method=="peakC.deseq2")$gene %>% unique(.)),
                         full.vec=(filter(gtf.df,gene_type=="protein_coding")$ensgene %>% unique(.))) # 76 genes; EF = 0.85; pval = 0.96

test1b <- gene_set_enrich(query.vec=(filter(gene.df)$gene %>% unique(.)),
                         full.vec=(filter(gtf.df)$ensgene %>% unique(.))) # 545 genes; EF = 0.92; pval=0.99
test2b <- gene_set_enrich(query.vec=(filter(gene.df,method=="peakC")$gene %>% unique(.)),
                         full.vec=(filter(gtf.df)$ensgene %>% unique(.))) # 135 genes; EF = 0.96; pval=0.70
test3b <- gene_set_enrich(query.vec=(filter(gene.df,method=="DESeq2")$gene %>% unique(.)),
                         full.vec=(filter(gtf.df)$ensgene %>% unique(.))) # 142 genes; EF = 0.99; pval = 0.54
test4b <- gene_set_enrich(query.vec=(filter(gene.df,method=="both")$gene %>% unique(.)),
                         full.vec=(filter(gtf.df)$ensgene %>% unique(.))) # 109 genes; EF = 0.94; pval = 0.76
test5b <- gene_set_enrich(query.vec=(filter(gene.df,method=="peakC.deseq2")$gene %>% unique(.)),
                         full.vec=(filter(gtf.df)$ensgene %>% unique(.))) # 159 genes; EF = 0.8; pval = 0.998

```


Summary Notes:

in total, all interactions implicate 545 genes through interactions within +2kb/-5kb of TSS, 275 (50.5%) of which are protein coding genes, the mean islet expression specificity score is 0.3590 for this set of 275 genes

Also, notably, PROTEIN-CODING genes uniquely impicated by DESeq2 only were significantly more likely to be specifically expressed in islet compared to adipose, liver, and muslce (ESS scores); enrichment factor = 1.2; pval = 0.023 




