---
title: "evaluate gene sets"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup 

```{r}

"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")


local.dir <- "/home/jason/science/projects/wtsa/joint_analyses/"
rescomp.dir <- "/home/jason/science/servers/FUSE5/"
work.dir <- rescomp.dir %&% "projects/wtsa/joint_analyses/"
file.dir <- work.dir %&% "analysis_files/"
grp.df  <- fread(file.dir %&% "all_locus_interactions_merge1K.txt")
gene.df <- fread(file.dir %&% "all_interaction_genes.txt")

```

Sanity check for consistency in gene counts between data frames 

```{r}

check.vec <- map(grp.df$genes,function(s){
  strsplit(s,split=",")[[1]]
}) %>% unlist(.) %>% as.character(.) %>% unique(.) 

sum(check.vec %in% gene.df$gene) # 545 
sum(gene.df$gene %in% check.vec) #545 

```

Now to separate interaction results for controls and targets among the enhancer captures (1st and 2nd rounds) 

```{r}

# Enhancer target captures
etarg.df <- dplyr::filter(grp.df,experiment=="enhancer.1st"|experiment=="enhancer.2nd") # 2,148 interactions corresponding to enhancer 1st or 2nd rounds 
econt.df <- filter(etarg.df,grepl("__uc__",capture) | grepl("__dc__",capture) ) # There are 288 controls involving control captures 
etarg.df <- filter(etarg.df,!(capture %in% econt.df$capture)) # There are 1,860 enhancer target capture interactions 

 <- etarg.df$genes %>% map(.,function(s){
  strsplit(s,split=",")[[1]]
}) %>% unlist(.) %>% as.character(.) %>% unique(.) # There are 432 Ensembl genes involved in interactions with enhancer targets captures
etarg.genes.peakc <- (filter(gene.df,gene %in% etarg.genes.all) %>% 
  filter(.,method=="peakC"|method=="both"|method=="peakC.deseq2"))$gene %>% unique(.) # 325 of the 432 genes are supported by peakC 
etarg.genes.deseq2 <- (filter(gene.df,gene %in% etarg.genes.all) %>% 
  filter(.,method=="DESeq2"|method=="both"))$gene %>% unique(.) # 200 of the 432 genes are supported by DESeq2 
etarg.genes.both <- (filter(gene.df,gene %in% etarg.genes.all) %>% 
  filter(.,method=="both"))$gene %>% unique(.) # 93 of the 432 genes are supported by peakC and DESeq2 

econt.genes.all <- econt.df$genes %>% map(.,function(s){
  strsplit(s,split=",")[[1]]
}) %>% unlist(.) %>% as.character(.) %>% unique(.) # There are 97 Ensembl genes involved in interactions with enhancer control captures
econt.genes.peakc <- (filter(gene.df,gene %in% econt.genes.all) %>% 
  filter(.,method=="peakC"|method=="both"|method=="peakC.deseq2"))$gene %>% unique(.) # 86 of the 97 genes are supported by peakC 
econt.genes.deseq2 <- (filter(gene.df,gene %in% econt.genes.all) %>% 
  filter(.,method=="DESeq2"|method=="both"))$gene %>% unique(.) # 48 of the 97 genes are supported by DESeq2 
econt.genes.both <- (filter(gene.df,gene %in% econt.genes.all) %>% 
  filter(.,method=="both"))$gene %>% unique(.) # 37 of the 97 genes are supported by peakC and DESeq2 

etarg.gene.list <- list(etarg.genes.all,etarg.genes.peakc,etarg.genes.deseq2,econt.genes.both)
econt.gene.list <- list(econt.genes.all,econt.genes.peakc,econt.genes.deseq2,econt.genes.both)

saveRDS(etarg.gene.list,file=file.dir %&% "enhancer.target.genes.RDS")
saveRDS(econt.gene.list,file=file.dir %&% "enhancer.control.genes.RDS")

```

# Gene-set enrichment test with SetRank

```{r}

library("rtracklayer")

ens.df <- fread(rescomp.dir %&% "datasets/Ensembl_HumanGenes_GRCh37-p13.txt")

gtf <- rtracklayer::import(rescomp.dir %&% "datasets/gencode.v19.genes.v7.patched_contigs.gtf")
gtf.df=as.data.frame(gtf)
pb <- txtProgressBar(min=0,max=dim(gtf.df)[1],style=3)
gtf.df$ensgene <- map(1:dim(gtf.df)[1],function(i){
  setTxtProgressBar(pb,i)
  ens <- gtf.df$gene_id[i]
  (strsplit(ens,split=".",fixed=TRUE))[[1]][1]
}) %>% as.character(.)

```


Summary: 

2,148 interactions corresponding to enhancer 1st or 2nd rounds 
There are 288 controls involving control captures
There are 1,860 enhancer target capture interactions 

There are 432 Ensembl genes involved in interactions with enhancer targets captures
325 of the 432 genes are supported by peakC 
200 of the 432 genes are supported by DESeq2 
93 of the 432 genes are supported by peakC and DESeq2

There are 97 Ensembl genes involved in interactions with enhancer control captures
86 of the 97 genes are supported by peakC
48 of the 97 genes are supported by DESeq2
37 of the 97 genes are supported by peakC and DESeq2
