---
title: "setRank"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup 

```{r}

"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")

local.dir <- "/home/jason/science/projects/wtsa/joint_analyses/"
rescomp.dir <- "/home/jason/science/servers/FUSE5/"
work.dir <- rescomp.dir %&% "projects/wtsa/joint_analyses/"
file.dir <- work.dir %&% "analysis_files/"
etarg.list  <- readRDS(file.dir %&% "enhancer.target.genes.RDS")
econt.list <- readRDS(file.dir %&% "enhancer.control.genes.RDS")

```


# Set Rank (Cedric Simillion)

## Build Annotation Tables 

Note that gene ids are in Entrez ids 

```{r}

library("SetRank")

db.pth <- "/home/jason/science/projects/wtsa/joint_analyses/GeneSets/package_data/" %&% "GeneSets.Homo.sapiens/data/"
biocyc.df <- fread(db.pth %&% "BIOCYC.txt")
gocc.df <- fread(db.pth %&% "GOCC.txt")
gobp.df <- fread(db.pth %&% "GOBP.txt")
gomf.df <- fread(db.pth %&% "GOMF.txt")
kegg.df <- fread(db.pth %&% "KEGG.txt")
reactome.df <- fread(db.pth %&% "REACTOME.txt")
pthway_inter_db.df <- fread(db.pth %&% "Pathway_Interaction_Database.txt")
wiki_pthwy.df <- fread(db.pth %&% "WikiPathways.txt")
    

```



## Selection of background genes (GTEx criteria used for gene selection prior to eQTL mapping)

Genes were selected based on expression thresholds of >0.1 TPM in at least 20% of samples and ≥6 reads in at least 20% of samples.



```{r}

tpm.df <- fread(work.dir %&% "islet-expression-data/01.03.2019.Islets_RNA-seq.gene.tpm.tsv")
count.df <- fread(work.dir %&% "islet-expression-data/01.03.2019.Islets_RNA-seq.gene.counts.tsv")
samp.df <- fread(work.dir %&% "islet-expression-data/sample_batches.txt",header=T) %>% 
  filter(.,!(grepl("Endo",sample))) %>% filter(.,!(grepl("_",sample))) %>% filter(.,!(grepl("ZMI",sample))) %>%
  filter(.,!(grepl("SCRAM",sample))) %>% filter(.,!(grepl("UN",sample))) %>% filter(.,!(grepl("GF",sample))) # 250 islet samples 

pb <- txtProgressBar(min=0,max=dim(tpm.df)[1],style=3)
tpm.sub  <- c()
for (i in 1:dim(tpm.df)[1]){
  setTxtProgressBar(pb,i)
  row.df <- tpm.df[i,] %>% dplyr::select(.,one_of("GeneID",samp.df$sample))
  tpm.vec <- row.df[,3:dim(row.df)[2]] %>% as.numeric(.)
  eval <- sum(tpm.vec > 0.1) > round(0.2 * length(tpm.vec))
  if (eval==TRUE){
    tpm.sub <- rbind(tpm.sub,row.df)
  }
}

pb <- txtProgressBar(min=0,max=dim(count.df)[1],style=3)
count.sub  <- c()
for (i in 1:dim(count.df)[1]){
  setTxtProgressBar(pb,i)
  row.df <- count.df[i,] %>% dplyr::select(.,one_of("GeneID",samp.df$sample))
  count.vec <- row.df[,3:dim(row.df)[2]] %>% as.numeric(.)
  eval <- sum(count.vec > 6) > round(0.2 * length(count.vec))
  if (eval==TRUE){
    count.sub <- rbind(count.sub,row.df)
  }
}

keep.genes <- unique(count.sub$GeneID[count.sub$GeneID %in% tpm.sub$GeneID]) # There are 22,929 genes to retain in islets that meet the filtering thresholds

```


Ensembl to Entrez gene conversion 

```{r}

x <- org.Hs.egENSEMBL
mapped_genes <- mappedkeys(x)
xx <- as.list(x[mapped_genes])
pb <- txtProgressBar(min=0,max=length(xx),style=3)
ens2ent.df <- c() 
for (i in 1:length(xx)){
  setTxtProgressBar(pb,i)
  entrez <- names(xx)[i]
  ens.vec <- xx[[i]]
  for (ens in ens.vec){
    build.df <- data.frame("entrez"=entrez,"ensembl"=ens,
                           stringsAsFactors = F)
    ens2ent.df <- rbind(ens2ent.df,build.df)
  }
}

```


```{r}

reference.genes <- map(keep.genes,function(s){
  ens <- strsplit(s,split=".",fixed=TRUE)[[1]][1]
  filter(ens2ent.df,ensembl==ens)$entrez %>% unique(.)
}) %>% unlist(.) %>% as.integer(.)

```


## Create gene set collection objects 

```{r}


collection.biocyc <- buildSetCollection(biocyc.df,referenceSet = reference.genes,
                                 maxSetSize = 500)
collection.gocc <- buildSetCollection(gocc.df,referenceSet = reference.genes,
                                 maxSetSize = 500)
collection.gobp <- buildSetCollection(gobp.df,referenceSet = reference.genes,
                                 maxSetSize = 500)
collection.gomf <- buildSetCollection(gomf.df,referenceSet = reference.genes,
                                 maxSetSize = 500)
collection.kegg <- buildSetCollection(kegg.df,referenceSet = reference.genes,
                                 maxSetSize = 500)
collection.reactome <- buildSetCollection(reactome.df,referenceSet = reference.genes,
                                 maxSetSize = 500)
collection.pthway_inter_db <- buildSetCollection(pthway_inter_db.df,referenceSet = reference.genes,
                                 maxSetSize = 500)
collection.wiki_pthwy <- buildSetCollection(wiki_pthwy.df,referenceSet = reference.genes,
                                 maxSetSize = 500)

```



## set rank analysis 

Test 

```{r}


options(mc.cores=4)

coll.list <- list(collection.gocc,collection.gobp,collection.gomf,collection.kegg,
                  collection.reactome,collection.biocyc,collection.pthway_inter_db,collection.wiki_pthwy)
coll.names <- c("GOCC","GOBP","GOMF","KEGG","REACTOME","biocyc","pathway_inter_db","wiki_pthwy")



etarg.names <- c("EnhTargets_All","EnhTargets_peakC","EnhTargets_DESeq2","EnhTargets_both")
for (i in 1:length(etarg.list)){
  gene.vec <- map(etarg.list[[i]],function(s){
    ens <- strsplit(s,split=".",fixed=TRUE)[[1]][1]
    filter(ens2ent.df,ensembl==ens)$entrez %>% unique(.)
  }) %>% unlist(.) %>% as.integer(.)
  etarg <- etarg.names[i]
  print(etarg)
  for (e in 1:length(coll.list)){
    coll <- coll.list[[e]]
    coll.name <- coll.names[e]
    print(coll.name)
    network <- setRankAnalysis(geneIDs = gene.vec, setCollection = coll,
                    use.ranks = FALSE,
                    setPCutoff = 0.01,fdrCutoff = 0.05)
    tryCatch(expr={exportSingleResult(network=network, selectedGenes=gene.vec, 
                          collection=coll, networkName = coll.name,
                          outputPath = file.dir%&%"SetRank." %&% etarg %&% "." %&% coll.name)},
             error=function(e){
               print("Error within SetRank, skipping")
             })
             
  }
}


```



```{r}

econt.names <- c("EnhControls_All","EnhControls_peakC","EnhControls_DESeq2","EnhControls_both")
for (i in 1:length(econt.list)){
  gene.vec <- map(econt.list[[i]],function(s){
    ens <- strsplit(s,split=".",fixed=TRUE)[[1]][1]
    filter(ens2ent.df,ensembl==ens)$entrez %>% unique(.)
  }) %>% unlist(.) %>% as.integer(.)
  econt <- econt.names[i]
  print(etarg)
  for (e in 1:length(coll.list)){
    coll <- coll.list[[e]]
    coll.name <- coll.names[e]
    print(coll.name)
    network <- setRankAnalysis(geneIDs = gene.vec, setCollection = coll,
                    use.ranks = FALSE,
                    setPCutoff = 0.01,fdrCutoff = 0.05)
    tryCatch(expr={exportSingleResult(network=network, selectedGenes=gene.vec, 
                          collection=coll, networkName = coll.name,
                          outputPath = file.dir%&%"SetRank." %&% econt %&% "." %&% coll.name)},
             error=function(e){
               print("Error within SetRank, skipping")
             })
             
  }
}

```

