---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup 

```{r}

"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")

local.dir <- "/home/jason/science/projects/wtsa/joint_analyses/"
rescomp.dir <- "/home/jason/science/servers/FUSE5/"
work.dir <- rescomp.dir %&% "projects/wtsa/joint_analyses/"
file.dir <- work.dir %&% "analysis_files/"
setrank.dir <- file.dir %&% "SetRank_output"

```



```{r}

filevec <- list.dirs(setrank.dir)
filevec <- filevec[grepl("Enh",filevec)]

namevec <- map(filevec, function(s){
  vec <- strsplit(s,split="/")[[1]]
  vec[length(vec)]
}) %>% as.character(.)


out.df <- c()
pb <- txtProgressBar(min=0,max=length(namevec),style=3)
for (i in 1:length(namevec)){
  setTxtProgressBar(pb,i)
  n <- namevec[i]
  vec <- strsplit(n,split=".",fixed=TRUE)[[1]]
  database <- vec[3]
  term2 <- vec[2]
  vec2 <- strsplit(x=term2,split="_")[[1]]
  capture_set <- vec2[1]
  caller <- vec2[2]
  fname <- setrank.dir %&% "/" %&% n %&% "/" %&% database %&% "_pathways.txt"
  if (file.exists(fname)==TRUE){
    df0 <- data.frame("capture_set"=capture_set,"caller"=caller,stringsAsFactors = FALSE)
    df <- fread(fname) 
    build.df <- cbind(df0,df)
    out.df <- rbind(out.df,build.df)
  }
}

```


```{r}

econt.df <- filter(out.df,capture_set=="EnhControls") %>% 
  arrange(.,caller,database,adjustedPValue) # 72 pathways enriched for control capture genes 
etarg.df <- filter(out.df,capture_set=="EnhTargets") %>% 
  arrange(.,caller,database,adjustedPValue) # 128 pathways enrihed for target capture genes  
filter(econt.df,caller=="All")$database %>% table(.)
#    GOBP     GOCC     GOMF     KEGG REACTOME 
#       9        4        3        1        4 
filter(etarg.df,caller=="All")$database %>% table(.)
#                        GOBP                         GOCC                         GOMF                         KEGG 
#                          21                            2                            6                            5 
#Pathway Interaction Database                     REACTOME                 WikiPathways 
#                           1                            5                            7 

write.table(x=econt.df,file=file.dir%&%"SetRank_output_enhancer-controls.txt",sep="\t",quote=F,row.names=F)
write.table(x=etarg.df,file=file.dir%&%"SetRank_output_enhancer-targets.txt",sep="\t",quote=F,row.names=F)

targ.vec <- filter(etarg.df,caller=="All")$description %>% unique(.)
cont.vec <- filter(econt.df,caller=="All")$description %>% unique(.)
targ.spec <- targ.vec[!(targ.vec %in% cont.vec)]

spec.df <- filter(etarg.df,caller=="All",description%in%targ.spec) %>% arrange(.,database,adjustedPValue) # 47 pathways 
write.table(x=spec.df,file=file.dir%&%"SetRank_output_enhancer-targets_targetSpecificPathways.txt",sep="\t",quote=F,row.names=F)

cont.spec <- cont.vec[!(cont.vec %in% targ.vec)] # 21 pathways 


```


72 pathways enriched for control capture genes, across all gene subsets  
128 pathways enrihed for target capture genes, across all gene subsets 

Dabase table for control pathways implicated by all genes implicated by all calling methods: 
#    GOBP     GOCC     GOMF     KEGG REACTOME 
#       9        4        3        1        4 
Dabase table for target pathways implicated by all genes implicated by all calling methods: 
#                        GOBP                         GOCC                         GOMF                         KEGG 
#                          21                            2                            6                            5 
#Pathway Interaction Database                     REACTOME                 WikiPathways 
#                           1                            5                            7 

47 of these pathways are specific to target captures 
21 pathways were specific to control captures 
