---
title: "4.4_OpenTargets-enrichment.Rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup 

```{r}

"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")

local.dir <- "/home/jason/science/projects/wtsa/joint_analyses/"
rescomp.dir <- "/home/jason/science/servers/FUSE5/"
work.dir <- rescomp.dir %&% "projects/wtsa/joint_analyses/"
file.dir <- work.dir %&% "analysis_files/"
etarg.list  <- readRDS(file.dir %&% "enhancer.target.genes.RDS")
econt.list <- readRDS(file.dir %&% "enhancer.control.genes.RDS")
keep.genes <- fread(file.dir %&% "islet-keep-genes.txt")$x
keep.genes <- map(keep.genes,function(g){
  strsplit(g,split=".",fixed=TRUE)[[1]][1]
}) %>% as.character(.) 

```

Subset gene sets to those meeting islet gene expression criteria (see 4.2 script) 

```{r}

set.seed(1)
etarg.all <- etarg.list[[1]][etarg.list[[1]] %in% keep.genes]
econt.all <- econt.list[[1]][econt.list[[1]] %in% keep.genes]
write.table(x=etarg.all,file=file.dir%&%"OpenTargets/"%&% "enhancer_target-genes_all.txt",
            sep="\t",quote=F,row.names=F,col.names=F)
write.table(x=econt.all,file=file.dir%&%"OpenTargets/"%&% "enhancer_control-genes_all.txt",
            sep="\t",quote=F,row.names=F,col.names=F)

etarg.all.subset <- etarg.all[etarg.all %in% econt.all]
set.seed(1)
etarg.samp <- sample(x=etarg.all[!(etarg.all %in% econt.all)],
                     size=(200-length(etarg.all.subset)),replace=FALSE)
etarg.subset <- c(etarg.samp,etarg.all.subset)
write.table(x=etarg.subset,file=file.dir%&%"OpenTargets/"%&% "enhancer_target-genes_all_random200.seed1.txt",
            sep="\t",quote=F,row.names=F,col.names=F)
```


Manually uploading gene list to OpenTargets Platform (https://www.targetvalidation.org/) for batch analyses:

CAUTION: The online batch query search only allows a maximum of 200 genes whereas, for all target capture genes, there was a total 273 (out of 432) Ensembl genes that meet islet expression criteria. So used a sample sample of 200 (keeping the 31 genes shared with controls) for enrichment, but will also do a manual annotation for all genes using the Open Targets Pyton API 

pip install opentargets

```{bash}

pip install --upgrade pip
pip install opentargets

```


```{python}

import sys
from opentargets import OpenTargetsClient
ot = OpenTargetsClient()
fin = open("/home/jason/science/servers/FUSE5/projects/wtsa/joint_analyses/analysis_files/OpenTargets/enhancer_target-genes_all.txt",'r')
fout = open("/home/jason/science/servers/FUSE5/projects/wtsa/joint_analyses/analysis_files/OpenTargets/enhancer_target-genes_all_OpenTarget-annotated.txt",'w')
head_list = ["gene","symbol","gene_Name","disease_ID","disease","overall_association_score","literature_score",
              "rna_expression_score","genetic_association_score","somatic_mutation_score","known_drug_score",
              "animal_model_score","affected_pathway_score","direct","drugs"]
fout.write("\t".join(head_list)+"\n")
count = 0
for line in fin:
  count += 1
  sys.stdout.write("\r%d"%count)
  sys.stdout.flush()  
  targ = line.strip()
  #print targ
  client = OpenTargetsClient()
  assoc = client.filter_associations()
  assoc.filter(target=str(targ),scorevalue_min=0.8,direct=True)
  #a_for_target = ot.get_associations_for_target(targ) # object comprised of dictionaries that can be indexed 
  #for a in a_for_target:
  for a in assoc:
    try:
      gene_symbol = a['target']['gene_info']['symbol']
      gene_full_name = a['target']['gene_info']['name']
      overall_association_score = a['association_score']['overall']
      literature_score = a['association_score']['datatypes']['literature']
      rna_expression_score = a['association_score']['datatypes']['rna_expression']
      genetic_association_score = a['association_score']['datatypes']['genetic_association']
      somatic_mutation_score = a['association_score']['datatypes']['somatic_mutation']
      known_drug_score = a['association_score']['datatypes']['known_drug']
      animal_model_score = a['association_score']['datatypes']['animal_model']
      affected_pathway_score = a['association_score']['datatypes']['affected_pathway']
      dis_id = a['disease']['id']
      #print dis_id
      disease = a['disease']['efo_info']['label']
      direct = a['is_direct'] # is the target directly or indirectly associated with the disease trait
      ## Search for specific evidence 
      #try: 
      #  client= OpenTargetsClient()
      #  response = client.filter_evidence()
      #  response.filter(target=str(targ),disease=str(dis_id))
      #  dic = {}
      #  for entry in response:
      #    mol_name = entry['drug']['molecule_name']
      #    mol_type = entry['drug']['molecule_type']
      #    try:
      #      dic[mol_name].append(mol_type)
      #    except:
      #      dic[mol_name] = [mol_type]
      #  for key in dic.keys():
      #    dic[key] = list(set(dic[key]))
      #  drugs = ",".join(dic.keys())
       # #drug_types = ",".join(dic.values()[0])
      #except:
      #  drugs = "NA"
      #  drug_types = "NA"
      write_list = [str(targ),str(gene_symbol),str(gene_full_name),str(dis_id),disease,str(overall_association_score),
                  str(literature_score),str(rna_expression_score),str(genetic_association_score),
                  str(somatic_mutation_score),str(known_drug_score),str(animal_model_score),
                  str(affected_pathway_score),str(direct)] #drugs]#,drug_types]
      fout.write("\t".join(write_list)+"\n")
    except:
      print "\nassociation failed\n"
    
fin.close()
fout.close()


```


```{r}

ot.df <- fread("/home/jason/science/servers/FUSE5/projects/wtsa/joint_analyses/analysis_files/OpenTargets/enhancer_target-genes_all_OpenTarget-annotated.txt") %>% dplyr::select(.,-one_of("drugs"))

count.df <- (ot.df$disease %>% table(.) %>% sort()) %>% as.data.frame(.)
names(count.df) <- c("Disease","Count")
count.df <- arrange(count.df,desc(Count))

write.table(x=ot.df,file=file.dir%&%"OpenTargets/"%&% "enhancer_target-genes_all_OpenTarget-annotated.txt",
            sep="\t",quote=F,row.names=F,col.names=T)
write.table(x=count.df,file=file.dir%&%"OpenTargets/"%&% "enhancer_target-genes_all_OpenTarget-diseaseCounts.txt",
            sep="\t",quote=F,row.names=F,col.names=T)
write.table(x=ot.df,file=file.dir%&%"OpenTargets/"%&% "enhancer_target-genes_all_OpenTarget-annotated.txt",
            sep="\t",quote=F,row.names=F,col.names=T)

diab.sub <- filter(ot.df,grepl("iabetes",disease)) # 101
diab.sub$gene %>% unique(.) %>% length(.) # 46 
write.table(x=diab.sub,file=file.dir%&%"OpenTargets/"%&% "enhancer_target-genes_all_OpenTarget-annotated_diabetes.txt",
            sep="\t",quote=F,row.names=F,col.names=T)

t2d.sub <- filter(ot.df,disease=="type II diabetes mellitus") # 38 
write.table(x=t2d.sub,file=file.dir%&%"OpenTargets/"%&% "enhancer_target-genes_all_OpenTarget-annotated_t2d.txt",
            sep="\t",quote=F,row.names=F,col.names=T)

ot.df$gene %>% unique(.) %>% length(.) # 170 

filter(diab.sub,genetic_association_score>0.8) %>% dim(.)

nongen.df <- filter(diab.sub,genetic_association_score<0.5) # 11 

```


There are 170 genes associated with disease outcomes (1,112) in OpenTarget platform at an overall association score > 0.8), 46 of which have been directly implicated in some form of diabetes and correspond to a total of 101 OpenTarget diabetes associations.

38 genes are associated with type 2 diabetes: 
[1] "GRB14"    "PPARG"    "ATXN7"    "PSMD6"    "ADCY5"    "FBXW7"    "NDUFS4"   "ARL15"    "PDE8B"    "RREB1"    "KCNK17"
[12] "KCNK16"   "JAZF1"    "GCK"      "MYL7"     "YKT6"     "TP53INP1" "DMRTA1"   "MTAP"     "ABO"      "GPSM1"    "DNLZ" 
[23] "ZMIZ1"    "HHEX"     "EXOC6"    "TCF7L2"   "CDKN1C"   "TRPM5"    "KCNJ11"   "NDUFS3"   "SPPL3"    "HNF1A" "C12orf43"
[34] "HMG20A"   "PEAK1"    "GIPR"     "EML2"     "HNF4A" 

Its notable that all but one of these genes have a genetic association score > 0.5. The exception is NDUFS3 (NADH:ubiquinone oxidoreductase core subunit S3) which only has a contribution from the known drug score to the overall association score. This gene is annotated as a target for the small molecule drug Metformin 

