---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup 


```{r}

"%&%" <- function(a,b) paste0(a,b)

library("tidyverse")
library("data.table")
library("GenomicRanges")

serv.dir <- "/home/jason/science/servers/FUSE5/" # Rescomp
proj.dir <- serv.dir %&% "projects/wtsa/enhancer_driven/second_round/"
work.dir <- proj.dir %&% "probe_design/"

cred.df <- fread(serv.dir%&%"projects/t2d_classification/revamp/genetic_credible_sets/gencred.txt")
atac.dir <- serv.dir %&% "projects/atac_analyses/evaluate_peaks/eLife2018/"
atac.df <- fread(atac.dir %&% "combined-cpm.txt")

atac.mat <- dplyr::select(atac.df[,7:24],-one_of("HP1507_UW")) %>% as.matrix(.)
peak_mean <- rowMeans(atac.mat)
atac.df$peak_mean <- peak_mean


```


Filter T2D credible SNPs (Mahajan et al 2018) with PPA >= 0.10

```{r}

sigcred.df <- filter(cred.df,PPA>=0.10)
sigcred05.df <- filter(cred.df,PPA>=0.05)
sigcred01.df <- filter(cred.df,PPA>=0.01)


```

Append ATAC information 

```{r}

append_atac <- function(df){
  atac <- c()
  peak_mean <- c()
  pb <- txtProgressBar(min=0,max=dim(df)[1],style=3)
  for (i in 1:dim(df)[1]){
    setTxtProgressBar(pb,i)
    c <- df$CHR[i];pos <- df$POS[i]
    sub <- filter(atac.df,Chr==c,Start<=pos,End>=pos)
    if (dim(sub)[1]==0){
      atac <- append(atac,FALSE)
      peak_mean <- append(peak_mean,NA)
    } else{
      atac <- append(atac,TRUE)
      peak_mean <- append(peak_mean,sub$peak_mean)      
    }
  }
  df$atac <- atac
  df$peak_mean <- peak_mean
  return(df)
}

```


```{r}

sigcred.df <- append_atac(sigcred.df)
sigcred05.df <- append_atac(sigcred05.df)
sigcred01.df <- append_atac(sigcred01.df)

```


# Cross-reference with First Round SNPs 


```{r}

assess_in_first <- function(sigatac.df){
  first.df <- fread(work.dir %&% "FirstRound-AllSNPsOfInterest.txt") %>%
    dplyr::select(one_of("Capture.Status","ID","Fragment"))
  first.df$CHR <- map(first.df$Fragment,function(s){
    strsplit(s,split=":")[[1]][1]
  }) %>% as.character(.)
  first.df$Frag.Start <- map(first.df$Fragment,function(s){
    (strsplit(s,split="-")[[1]][1] %>% strsplit(.,split=":"))[[1]][2]
  }) %>% as.integer(.)
  first.df$Frag.End <- map(first.df$Fragment,function(s){
    strsplit(s,split="-")[[1]][2]
  }) %>% as.integer(.)
  first.round <- map(1:dim(sigatac.df)[1],function(i){
    sub <- sigatac.df[i,]
    c <- sub$CHR; pos <- sub$POS
    sub2 <- filter(first.df,CHR==c,Frag.Start<=pos,Frag.End>=pos)
    ifelse(dim(sub2)[1]==0,FALSE,ifelse(sub2$Capture.Status=="HighQuality.paired",TRUE,FALSE))
  }) %>% as.logical(.)
  fr.id <- map(1:dim(sigatac.df)[1],function(i){
    sub <- sigatac.df[i,]
    c <- sub$CHR; pos <- sub$POS
    sub2 <- filter(first.df,CHR==c,Frag.Start<=pos,Frag.End>=pos)
    ifelse(dim(sub2)[1]==0,NA,ifelse(sub2$Capture.Status=="HighQuality.paired",sub2$ID,NA))
  }) %>% as.character(.)
  sigatac.df$first.round <- first.round
  sigatac.df$fr.id <- fr.id  
  return(sigatac.df)
}

```



Filter to those with ATAC overlap 

```{r}

sigatac.df <- sigcred.df %>% filter(.,atac==TRUE) %>% assess_in_first(.)
sigatac05.df <- sigcred05.df %>% filter(.,atac==TRUE) %>% assess_in_first(.)
sigatac01.df <- sigcred01.df %>% filter(.,atac==TRUE) %>% assess_in_first(.)

```

There are 159 credible SNPs (from 120 loci) that have PPA >=0.10 and overlap islet ATAC peaks (MACS2 on n=17 islet ATAC samples) using filtered ATAC peaks 

There are 257 credible SNPs (from 167 loci) that have PPA >=0.10 and overlap islet ATAC peaks (MACS2 on n=17 islet ATAC samples) using filtered ATAC peaks 

There are 665 credible SNPs (from 280 loci) that have PPA >=0.01 and overlap islet ATAC peaks (MACS2 on n=17 islet ATAC samples) using filtered ATAC peaks




```{r}

new.df <- filter(sigatac.df,first.round==FALSE)
write.table(x=new.df,file=work.dir %&% "second-round_T2D-SNPs_PPA10.txt",sep="\t",quote=F,row.names=F)
write.table(x=sigatac.df,file=work.dir %&% "second-round_T2D-SNPs_PPA05_withFirstRound.txt",sep="\t",quote=F,row.names=F)

new05.df <- filter(sigatac.df,first.round==FALSE)
write.table(x=new05.df,file=work.dir %&% "second-round_T2D-SNPs_PPA05.txt",sep="\t",quote=F,row.names=F)
write.table(x=sigatac05.df,file=work.dir %&% "second-round_T2D-SNPs_PPA05_withFirstRound.txt",sep="\t",quote=F,row.names=F)

new01.df <- filter(sigatac01.df,first.round==FALSE)
write.table(x=new01.df,file=work.dir %&% "second-round_T2D-SNPs_PPA01.txt",sep="\t",quote=F,row.names=F)
write.table(x=sigatac01.df,file=work.dir %&% "second-round_T2D-SNPs_PPA01_withFirstRound.txt",sep="\t",quote=F,row.names=F)
```


First Round variants 

```{r}

#first.df <- filter(sigatac.df,first.round==TRUE)
#first.df$Locus.ID %>% unique(.) %>% length(.)
#first.df$symbol %>% unique(.)
#write.table(x=first.df,file=work.dir %&% "second-round_T2D-SNPs_PPA10_OnlyInFirstRound.txt",sep="\t",quote=F,row.names=F)


```

