---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup 


```{r}

"%&%" <- function(a,b) paste0(a,b)

library("tidyverse")
library("data.table")
library("GenomicRanges")

serv.dir <- "/home/jason/science/servers/FUSE5/" # Rescomp
serv.dir2 <- "/home/jason/science/servers/FUSE2/" # CBRG

proj.dir <- serv.dir %&% "projects/wtsa/enhancer_driven/second_round/"
work.dir <- proj.dir %&% "probe_design/"


gt.df <- fread(work.dir %&% "second-round_ENGAGE-SNPs_PPA01_withFirstRound.txt")
t2d.df <- fread(work.dir %&% "second-round_T2D-SNPs_PPA01_withFirstRound.txt")


```


# Combine DF 

```{r}

snps <- c(gt.df$SNPID,t2d.df$SNPID) %>% unique(.)

df <- c()
pb <- txtProgressBar(min=0,max=length(snps),style=3)
for (i in 1:length(snps)){
  setTxtProgressBar(pb,i)
  s <- snps[i]
  a <- filter(t2d.df,SNPID==s) %>% select(.,one_of("SNPID","CHR","POS","peak_mean","first.round"))
  b <- filter(gt.df,SNPID==s) %>% select(.,one_of("SNPID","CHR","POS","peak_mean","first.round"))
  c <- rbind(a,b); c <- c[!duplicated(c),]
  df <- rbind(df,c)
}

df$t2d <- map(1:(dim(df)[1]),function(i){
  s <- df$SNPID[i]
  sub <- filter(t2d.df,SNPID==s)
  ifelse(dim(sub)[1]>0,TRUE,FALSE)
}) %>% as.logical(.)
df$bmi <- map(1:(dim(df)[1]),function(i){
  s <- df$SNPID[i]
  sub <- filter(gt.df,SNPID==s,trait=="BMI")
  ifelse(dim(sub)[1]>0,TRUE,FALSE)
}) %>% as.logical(.)
df$fg <- map(1:(dim(df)[1]),function(i){
  s <- df$SNPID[i]
  sub <- filter(gt.df,SNPID==s,trait=="FG")
  ifelse(dim(sub)[1]>0,TRUE,FALSE)
}) %>% as.logical(.)
df$fi <- map(1:(dim(df)[1]),function(i){
  s <- df$SNPID[i]
  sub <- filter(gt.df,SNPID==s,trait=="FI")
  ifelse(dim(sub)[1]>0,TRUE,FALSE)
}) %>% as.logical(.)


df$t2d.ppa <- map(1:(dim(df)[1]),function(i){
  s <- df$SNPID[i]
  sub <- filter(t2d.df,SNPID==s)$PPA
  ifelse(length(sub)>0,sub,NA)
}) %>% as.numeric(.)
df$bmi.ppa <- map(1:(dim(df)[1]),function(i){
  s <- df$SNPID[i]
  sub <- filter(gt.df,SNPID==s,trait=="BMI")$PPA
  ifelse(length(sub)>0,sub,NA)
}) %>% as.numeric(.)
df$fg.ppa <- map(1:(dim(df)[1]),function(i){
  s <- df$SNPID[i]
  sub <- filter(gt.df,SNPID==s,trait=="FG")$PPA
  ifelse(length(sub)>0,sub,NA)
}) %>% as.numeric(.)
df$fi.ppa <- map(1:(dim(df)[1]),function(i){
  s <- df$SNPID[i]
  sub <- filter(gt.df,SNPID==s,trait=="FI")$PPA
  ifelse(length(sub)>0,sub,NA)
}) %>% as.numeric(.)


```


Filter out SNPs that are only associated with BMI and/or FI 

```{r}

df.full <- df
com.df <- filter(df,t2d==TRUE | fg==TRUE)

```

There are 670 SNPs that have been fine-mapped to have PPA >= 0.01 for association with T2D and/or FG that also overlap with ATAC peaks (n=17) 


Determined if successfully captured in the first round 

```{r}

cbrg.dir <- serv.dir2 %&% "wtsa/enhancer_driven/first_round/"
failed.vec <- fread(cbrg.dir %&% "failed_captures_list_alt.txt",header=F)$V1
oligo.df <- fread(cbrg.dir %&% "newOligofile_shortNames_modified.txt",header=F)
success.df <- filter(oligo.df,!(V1 %in% failed.vec))
success.df$CHR <- "chr" %&% success.df$V2

com.df$captured.in.first <- map(1:dim(com.df)[1],function(i){
 row.df <- com.df[i,] 
 c <- row.df$CHR
 p <- row.df$POS
 sub <- filter(success.df,CHR==c,V3<=p,V4>=p)
 ifelse(dim(sub)[1]>0,TRUE,FALSE)
}) %>% as.logical(.)

new.df <- filter(com.df,captured.in.first==FALSE)

new.df$locus <- map(1:dim(new.df)[1],function(i){
  s <- new.df$SNPID[i]
  a <- filter(t2d.df,SNPID==s)$symbol
  b <- filter(gt.df,SNPID==s)$symbol
  ifelse(length(a)>0,a,b)
}) %>% as.character(.)
new.df$locus[25:26] <- "GCK"

(new.df$locus) %>% unique(.) %>% length(.)

```

Of the 670 SNPs (that have been fine-mapped to have PPA >= 0.01 for association with T2D and/or FG that also overlap with ATAC peaks (n=17), 624 SNPs (93.1%) at 188 loci have NOT BEEN CAPTURED in the first round 


```{r}

new05.df <- filter(new.df,t2d.ppa>=0.05 | fg.ppa>0.05)
(new05.df$locus) %>% unique(.) %>% length(.)

write.table(x=new05.df,work.dir %&% "second-round-credible-SNPs-to-capture.txt",sep="\t",quote=F,row.names=F)

```


Of the 670 SNPs (that have been fine-mapped to have PPA >= 0.01 for association with T2D and/or FG that also overlap with ATAC peaks (n=17), 220 SNPs (32.8%) at 112 loci have NOT BEEN CAPTURED in the first round AND have a PPA >= 0.05 for T2D and/or FG 
