---
title: "evaluate-promoter-interactions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup 

```{r}
"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")
library("GenomicRanges")
rescomp.dir <- "/Users/jasont/science/servers/FUSE5/"
work.dir <- rescomp.dir %&% "projects/wtsa/joint_analyses/01_interaction-mapping/"
output.dir <- work.dir %&% "output_files/"
```

peakC interactions
```{r}
pkc.df1 <- fread(work.dir %&% "peakC_interactions/" %&%
  "promoter_endo_peakC-interactions.txt")
pkc.df2 <- fread(work.dir %&% "peakC_interactions/" %&%
  "promoter_endo_peakC-interactions_merge1K.txt")
```

deseq2 interactions 

```{r}
dsq.df1 <- fread(work.dir %&% "DESeq2_interactions/" %&%
  "promoter_endo_DESeq2-interactions.txt")
dsq.df2 <- fread(work.dir %&% "DESeq2_interactions/" %&%
  "promoter_endo_DESeq2-interactions_merge1K.txt")
```

peaky interactions
```{r}
pky.all.df1a <- fread(work.dir %&% "peaky_interactions/" %&%
  "promoter_endo_peaky-interactions_rjmcmc-pos01.txt")
pky.all.df2a <- fread(work.dir %&% "peaky_interactions/" %&%
  "promoter_endo_peaky-interactions_merge1K_rjmcmc-pos01.txt")
pky.all.df1b <- fread(work.dir %&% "peaky_interactions/" %&%
  "promoter_endo_peaky-interactions_rjmcmc-pos05.txt")
pky.all.df2b <- fread(work.dir %&% "peaky_interactions/" %&%
  "promoter_endo_peaky-interactions_merge1K_rjmcmc-pos05.txt")
pky.all.df1c <- fread(work.dir %&% "peaky_interactions/" %&%
  "promoter_endo_peaky-interactions_rjmcmc-pos10.txt")
pky.all.df2c <- fread(work.dir %&% "peaky_interactions/" %&%
  "promoter_endo_peaky-interactions_merge1K_rjmcmc-pos10.txt")
```

## Evaluate interactions 

### Interaction set 1: Merged only adjacent interactions 

```{r}
dim(pkc.df1)[1]
dim(dsq.df1)[1]
dim(pky.all.df1a)[1] 
dim(pky.all.df1b)[1]
dim(pky.all.df1c)[1]
# Note; peaky results in between 7-50X more interactions than other methods, but chromosome wide!
```

**Note: PeakC and DESeq2 interactions were called w/n 1Mb range, whereas peaky didn't have a range limit, will restrict peaky interactions to those w/n 1Mb** 

```{r}
pky.df1a <- filter(pky.all.df1a,abs(dist.from.vp) <= 1e6)
dim(pky.df1a)[1]
dim(pky.df1a)[1]/dim(pkc.df1)[1]
pky.df1b <- filter(pky.all.df1b,abs(dist.from.vp) <= 1e6)
dim(pky.df1b)[1]
dim(pky.df1b)[1]/dim(pkc.df1)[1]
pky.df1c <- filter(pky.all.df1c,abs(dist.from.vp) <= 1e6)
dim(pky.df1c)[1]
dim(pky.df1c)[1]/dim(pkc.df1)[1]
```

**Note: There are much more peaky interactions than peakC interactions when limited to 1Mb distance, about 3.3X-13.8X more interaction, depending on the rjmcmc_pos threshold used; 1557 (rjmcmc>=0.01), 626 (rjmcmc>=0.05), 374 (rjmcmc>=0.10) vs 113 (peakC) and 253 (deseq2) ** 

Will now compare the distribution of interaction widths across the three methods

```{r}
summary(pkc.df1$width)
summary(dsq.df1$width)
summary(pky.df1a$width)
summary(pky.df1b$width)
summary(pky.df1c$width)

median(pkc.df1$width)
median(dsq.df1$width)
median(pky.df1a$width)
median(pky.df1b$width)
median(pky.df1c$width)
```

**Note: peakC interactions are much wider than peaky/deseq2 interactions, by a factor of about 10X; 2.956Kb vs 337bp-381bp(peaky) vs 326bp**

Will now evaluate distribution of absolute distance to vp 

```{r}
summary(pkc.df1$dist.from.vp %>% abs(.))
summary(dsq.df1$dist.from.vp %>% abs(.))
summary(pky.df1a$dist.from.vp %>% abs(.))
summary(pky.df1b$dist.from.vp %>% abs(.))
summary(pky.df1c$dist.from.vp %>% abs(.))

median(pkc.df1$dist.from.vp %>% abs(.))
median(dsq.df1$dist.from.vp %>% abs(.))
median(pky.df1a$dist.from.vp %>% abs(.))
median(pky.df1b$dist.from.vp %>% abs(.))
median(pky.df1c$dist.from.vp %>% abs(.))
```

**Note: There seems to be a clear distinction between the distance of these interactions to the vp across these methods; DESeq2 interactions are definitely closer to the viewpoint (median ~14Kb) whereas the model-based approaches are more distal with peakC having a median distance of 140Kb and peaky being 267Kb-297Kb (10X vs 19X-21X greater distances)**

#### Determine extent of overlap between interactions 

```{r}
pkc.gr1 <- GRanges(seqnames = pkc.df1$seqnames,
                   IRanges(start=pkc.df1$start,end=pkc.df1$end))
dsq.gr1 <- GRanges(seqnames = dsq.df1$seqnames,
                   IRanges(start=dsq.df1$start,end=dsq.df1$end))
pky.gr1a <- GRanges(seqnames = pky.df1a$seqnames,
                   IRanges(start=pky.df1a$start,end=pky.df1a$end))
pky.gr1b <- GRanges(seqnames = pky.df1b$seqnames,
                   IRanges(start=pky.df1b$start,end=pky.df1b$end))
pky.gr1c <- GRanges(seqnames = pky.df1c$seqnames,
                   IRanges(start=pky.df1c$start,end=pky.df1c$end))
```

Number of peakC interactions that overlap peaky set 
```{r}
sum(pkc.gr1 %over% pky.gr1a) # 58
sum(pkc.gr1 %over% pky.gr1a)/length(pkc.gr1) * 100 # 51.3% 

sum(pkc.gr1 %over% pky.gr1b) # 49
sum(pkc.gr1 %over% pky.gr1b)/length(pkc.gr1) * 100 # 43.4% 

sum(pkc.gr1 %over% pky.gr1c) # 45
sum(pkc.gr1 %over% pky.gr1c)/length(pkc.gr1) * 100 # 39.8% 
```

Number of peakC interactions that overlap DESeq2 set 
```{r}
sum(pkc.gr1 %over% dsq.gr1) # 20
sum(pkc.gr1 %over% dsq.gr1)/length(pkc.gr1) * 100 # 17.7% 
```

Number of Deseq interactions that overlap peaky set 
```{r}
sum(dsq.gr1 %over% pky.gr1a) # 65
sum(dsq.gr1 %over% pky.gr1a)/length(dsq.gr1) * 100 # 25.7% 

sum(dsq.gr1 %over% pky.gr1b) # 28
sum(dsq.gr1 %over% pky.gr1b)/length(dsq.gr1) * 100 # 11.1% 

sum(dsq.gr1 %over% pky.gr1c) # 23
sum(dsq.gr1 %over% pky.gr1c)/length(dsq.gr1) * 100 # 9.1% 
```

Number of DESeq2 interactions that overlap peakC set 
```{r}
sum(dsq.gr1 %over% pkc.gr1) # 39
sum(dsq.gr1 %over% pkc.gr1)/length(dsq.gr1) * 100 # 15.4% 
```

Number of peaky interactions that overlap DESeq2 set 
```{r}
sum(pky.gr1a %over% dsq.gr1) # 74
sum(pky.gr1a %over% dsq.gr1)/length(pky.gr1a) * 100 # 4.8% 

sum(pky.gr1b %over% dsq.gr1) # 28
sum(pky.gr1b %over% dsq.gr1)/length(pky.gr1b) * 100 # 4.5% 

sum(pky.gr1c %over% dsq.gr1) # 21
sum(pky.gr1c %over% dsq.gr1)/length(pky.gr1c) * 100 # 5.6% 
```

Number of peaky interactions that overlap peakC set 
```{r}
sum(pky.gr1a %over% pkc.gr1) # 112
sum(pky.gr1a %over% pkc.gr1)/length(pky.gr1a) * 100 # 7.2% 

sum(pky.gr1b %over% pkc.gr1) # 84
sum(pky.gr1b %over% pkc.gr1)/length(pky.gr1b) * 100 # 13.4% 

sum(pky.gr1c %over% pkc.gr1) # 62
sum(pky.gr1c %over% pkc.gr1)/length(pky.gr1c) * 100 # 16.6% 
```



### Interaction set 2: Merged interactions withn 1Kb of each other 

```{r}
dim(pkc.df2)[1]
dim(dsq.df2)[1]
dim(pky.all.df2a)[1] 
dim(pky.all.df2b)[1] 
dim(pky.all.df2c)[1] 
# Note; peaky rjmcmc_pos>=0.10 interactions about 5X-8X more than other methods; peaky rjmcmc_pos>=0.01 interactions about 33-55X more interactions than other methods, but chromosome wide! 
```

**Note: PeakC and DESeq2 interactions were called w/n 1Mb range, whereas peaky didn't have a range limit, will restrict peaky interactions to those w/n 1Mb** 

```{r}
pky.df2a <- filter(pky.all.df2a,abs(dist.from.vp) <= 1e6)
dim(pky.df2a)[1]
dim(pky.df2a)[1]/dim(pkc.df2)[1]

pky.df2b <- filter(pky.all.df2b,abs(dist.from.vp) <= 1e6)
dim(pky.df2b)[1]
dim(pky.df2b)[1]/dim(pkc.df2)[1]

pky.df2c <- filter(pky.all.df2c,abs(dist.from.vp) <= 1e6)
dim(pky.df2c)[1]
dim(pky.df2c)[1]/dim(pkc.df2)[1]
```

**Note: There are still more peaky interactions than peakC interactions when limited to 1Mb distance, about 3.7X-11.4X more interaction, depending on the rjmcmc_pos threshold used; 915 (rjmcmc>=0.01), 460 (rjmcmc>=0.05), 296 (rjmcmc>=0.10) vs 80 (peakC) and 131 (deseq2) ** 

Will now compare the distribution of interaction widths across the three methods

```{r}
summary(pkc.df2$width)
summary(dsq.df2$width)
summary(pky.df2a$width)
summary(pky.df2b$width)
summary(pky.df2c$width)

median(pkc.df2$width)
median(dsq.df2$width)
median(pky.df2a$width)
median(pky.df2b$width)
median(pky.df2c$width)
```

**Note: peakC interactions are  wider than peaky/deseq2 interactions, by a factor of about 2.6X; median 6.7Kb vs 2.6Kb (deseq) vs 2.6Kb-2.7Kb (peaky)**

Will now evaluate distribution of absolute distance to vp 

```{r}
summary(pkc.df2$dist.from.vp %>% abs(.))
summary(dsq.df2$dist.from.vp %>% abs(.))
summary(pky.df2a$dist.from.vp %>% abs(.))
summary(pky.df2b$dist.from.vp %>% abs(.))
summary(pky.df2c$dist.from.vp %>% abs(.))

median(pkc.df2$dist.from.vp %>% abs(.))
median(dsq.df2$dist.from.vp %>% abs(.))
median(pky.df2a$dist.from.vp %>% abs(.))
median(pky.df2b$dist.from.vp %>% abs(.))
median(pky.df2c$dist.from.vp %>% abs(.))
```

**Note: There seems to be a clear distinction between the distance of these interactions to the vp across these methods; DESeq2 interactions are definitely closer to the viewpoint (median ~22.4Kb) whereas the model-based approaches are more distal with peakC having a median distance of 124Kb and peaky being 267Kb-317Kb (5.5X vs 11.9X-14.2X greater distances)**


#### Determine extent of overlap between interactions 

```{r}
pkc.gr2 <- GRanges(seqnames = pkc.df2$seqnames,
                   IRanges(start=pkc.df2$start,end=pkc.df2$end))
dsq.gr2 <- GRanges(seqnames = dsq.df2$seqnames,
                   IRanges(start=dsq.df2$start,end=dsq.df2$end))
pky.gr2a <- GRanges(seqnames = pky.df2a$seqnames,
                   IRanges(start=pky.df2a$start,end=pky.df2a$end))
pky.gr2b <- GRanges(seqnames = pky.df2b$seqnames,
                   IRanges(start=pky.df2b$start,end=pky.df2b$end))
pky.gr2c <- GRanges(seqnames = pky.df2c$seqnames,
                   IRanges(start=pky.df2c$start,end=pky.df2c$end))
```

Number of peakC interactions that overlap peaky set 
```{r}
sum(pkc.gr2 %over% pky.gr2a) # 59
sum(pkc.gr2 %over% pky.gr2a)/length(pkc.gr2) * 100 # 73.8% 

sum(pkc.gr2 %over% pky.gr2b) #  50
sum(pkc.gr2 %over% pky.gr2b)/length(pkc.gr2) * 100 # 62.5% 

sum(pkc.gr2 %over% pky.gr2c) # 46
sum(pkc.gr2 %over% pky.gr2c)/length(pkc.gr2) * 100 # 57.5% 
```

Number of peakC interactions that overlap DESeq2 set 
```{r}
sum(pkc.gr2 %over% dsq.gr2) # 20
sum(pkc.gr2 %over% dsq.gr2)/length(pkc.gr2) * 100 # 25% 
```

Number of Deseq interactions that overlap peaky set 
```{r}
sum(dsq.gr2 %over% pky.gr2a) # 45
sum(dsq.gr2 %over% pky.gr2a)/length(dsq.gr2) * 100 # 34.4% 

sum(dsq.gr2 %over% pky.gr2b) # 31
sum(dsq.gr2 %over% pky.gr2b)/length(dsq.gr2) * 100 # 23.7% 

sum(dsq.gr2 %over% pky.gr2c) # 22
sum(dsq.gr2 %over% pky.gr2c)/length(dsq.gr2) * 100 # 16.8% 
```

Number of DESeq2 interactions that overlap peakC set 
```{r}
sum(dsq.gr2 %over% pkc.gr2) # 24
sum(dsq.gr2 %over% pkc.gr2)/length(dsq.gr2) * 100 # 18.3% 
```

Number of peaky interactions that overlap DESeq2 set 
```{r}
sum(pky.gr2a %over% dsq.gr2) # 58
sum(pky.gr2a %over% dsq.gr2)/length(pky.gr2a) * 100 # 6.3% 

sum(pky.gr2b %over% dsq.gr2) # 29
sum(pky.gr2b %over% dsq.gr2)/length(pky.gr2b) * 100 # 6.3% 

sum(pky.gr2c %over% dsq.gr2) # 23
sum(pky.gr2c %over% dsq.gr2)/length(pky.gr2c) * 100 # 7.7% 
```

Number of peaky interactions that overlap peakC set 
```{r}
sum(pky.gr2a %over% pkc.gr2) # 94
sum(pky.gr2a %over% pkc.gr2)/length(pky.gr2a) * 100 # 10.3% 

sum(pky.gr2b %over% pkc.gr2) # 76
sum(pky.gr2b %over% pkc.gr2)/length(pky.gr2b) * 100 # 16.5% 

sum(pky.gr2c %over% pkc.gr2) # 69
sum(pky.gr2c %over% pkc.gr2)/length(pky.gr2c) * 100 # 23.3% 
```

**Note: the overlap of peakC interactions among peaky interactions is much higher with the 1K merge extension 58%-74% vs 40%-51%; overall it seems that peaky overlaps shares of peakC and DEseq2 interactions, but not vice-versa, largely due to the greater number of peaky interactions, but a large share of these interactions overlap those for the other two methods** 


