---
title: "Untitled"
author: "Jason Torres"
date: "5/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup 

```{r}

"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")
library("ggbio")
library("RColorBrewer")
library("viridis")
library(Homo.sapiens)

rescomp.dir <- "/Users/jtorres/FUSE5/" #"/home/jason/science/servers/FUSE5/"
cbrg.dir <- "/Users/jtorres/FUSE2/" #"/home/jason/science/servers/FUSE2/"
got2d.dir <- "/Users/jtorres/FUSE/" #"/home/jason/science/servers/FUSE/" 

enh1.dir <- rescomp.dir %&% "projects/wtsa/enhancer_driven/first_round/statistical_anlayses/output_files/"
prom.dir <- rescomp.dir %&% "projects/wtsa/promoter_driven/statistical_analysis/output_files/"
enh1.df <- fread(enh1.dir %&% "peakC-interactions-default_withBaits.txt")
prom.df <- fread(prom.dir %&% "peakC-interactions-default_withBaits.txt")

```


# Profile interaction features 

```{r}

library("GenomicRanges")

build_interaction_df <- function(df){
  names(df)[1] <- "capture"
  out.df <- c()
  sub <- filter(df,type=="bait")
  for (i in 1:dim(sub)[1]){
    row.df <- sub[i,1:5]
    names(row.df) <- c("capture","chrom","bait.start","bait.end","bait.width")
    cap <- row.df$capture
    sub2 <- filter(df,capture==cap,type=="interaction")
    for (i in 1:dim(sub2)[1]){
      row.df2 <- sub2[i,3:5]
      names(row.df2) <- c("inter.start","inter.end","inter.width")
      r.df <- cbind(row.df,row.df2)
      out.df <- rbind(out.df,r.df)
    }
  }
  inter.dist <- map(1:dim(out.df)[1],function(i){
    row.df <- out.df[i,]
    bait.gr <- GRanges(seqnames = row.df$chrom,IRanges(row.df$bait.start,row.df$bait.end))
    inter.gr <- GRanges(seqnames = row.df$chrom,IRanges(row.df$inter.start,row.df$inter.end))
    distance(bait.gr,inter.gr)
  }) %>% as.integer(.)
  out.df$inter.dist <- inter.dist
  return(out.df)
}


build_summary_df <- function(df){
  int.df <- build_interaction_df(df)
  #num.unique.cap <- int.df$capture %>% unique(.) %>% length(.) # Number of captures w/ significant peakC interactions 
  cap.vec <- int.df$capture %>% unique(.)
  out.df <- c()
  for (cap in cap.vec){
    sub <- filter(int.df,capture==cap)
    build.df <- data.frame("capture"=cap,"chrom"=sub$chrom[1],"num.inter"=dim(sub)[1],
                           "med.inter.width"=median(sub$inter.width), "min.inter.dist"=min(sub$inter.dist),
                           "med.inter.dist"=median(sub$inter.dist),"max.inter.dist"=max(sub$inter.dist),stringsAsFactors = F)
    out.df <- rbind(out.df,build.df)
  }
  return(out.df)
}


```



# Promoter capture 

```{r}

sum.df1 <- build_summary_df(prom.df) 
summary(sum.df1$num.inter)

```

There are `r dim(sum.df1)[1]` captures with at least one significant peakC interaction 

The total number of peakC interactions from this experiment is  `r sum.df1$num.inter %>% sum(.)`

The median number of interactions at a capture site is `r median(sum.df1$num.inter)` 
The minimum number of interactions at a capture site is `r min(sum.df1$num.inter)` and correponds to `r arrange(sum.df1,num.inter)[1,]$capture` 
The maximum number of interactions at a capture site is `r max(sum.df1$num.inter)` and correponds to `r arrange(sum.df1,-(num.inter))[1,]$capture` 

The median interaction width at a capture site is `r median(sum.df1$med.inter.width)` 
The minimum interaction width at a capture site is `r min(sum.df1$med.inter.width)` and correponds to `r arrange(sum.df1,med.inter.width)[1,]$capture`
The maximum interaction width at a capture site is `r max(sum.df1$med.inter.width)` and correponds to `r arrange(sum.df1,-med.inter.width)[1,]$capture`

The median interaction distance at a capture site is `r median(sum.df1$med.inter.dist)` 
The minimum interaction distance at a capture site is `r min(sum.df1$min.inter.dist)` and correponds to `r arrange(sum.df1,min.inter.dist)[1,]$capture`
The maximum interaction distance at a capture site is `r max(sum.df1$max.inter.dist)` and correponds to `r arrange(sum.df1,-max.inter.dist)[1,]$capture`




# Enhancer capture (1st)

```{r}

sum.df2 <- build_summary_df(enh1.df) 
summary(sum.df2$num.inter)

```

There are `r dim(sum.df2)[1]` captures with at least one significant peakC interaction 

The total number of peakC interactions from this experiment is  `r sum.df2$num.inter %>% sum(.)`

The median number of interactions at a capture site is `r median(sum.df2$num.inter)` 
The minimum number of interactions at a capture site is `r min(sum.df2$num.inter)` and correponds to `r arrange(sum.df2,num.inter)[1,]$capture` 
The maximum number of interactions at a capture site is `r max(sum.df2$num.inter)` and correponds to `r arrange(sum.df2,-(num.inter))[1,]$capture` 

The median interaction width at a capture site is `r median(sum.df2$med.inter.width)` 
The minimum interaction width at a capture site is `r min(sum.df2$med.inter.width)` and correponds to `r arrange(sum.df2,med.inter.width)[1,]$capture`
The maximum interaction width at a capture site is `r max(sum.df2$med.inter.width)` and correponds to `r arrange(sum.df2,-med.inter.width)[1,]$capture`

The median interaction distance at a capture site is `r median(sum.df2$med.inter.dist)` 
The minimum interaction distance at a capture site is `r min(sum.df2$min.inter.dist)` and correponds to `r arrange(sum.df2,min.inter.dist)[1,]$capture`
The maximum interaction distance at a capture site is `r max(sum.df2$max.inter.dist)` and correponds to `r arrange(sum.df2,-max.inter.dist)[1,]$capture`


# Check for interaction overlaps between Promoter and Enhancer (1st Round) captures


```{r}

int.df1 <- build_interaction_df(prom.df)
int.df2 <- build_interaction_df(enh1.df)


build_overlap_check_df <- function(int.df1,int.df2){
  cvar.gr <- GRanges(int.df2$chrom,IRanges(int.df2$bait.start,int.df2$bait.end))
  out.df <- c()
  for (cap in unique(int.df1$capture)){
    sub.df <- filter(int.df1,capture==cap)
    for (i in 1:dim(sub.df)[1]){
      row.df <- sub.df[i,]
      prom.gr <- GRanges(seqnames=row.df$chrom,IRanges(row.df$bait.start,row.df$bait.end))
      promint.gr <- GRanges(seqnames=row.df$chrom,IRanges(row.df$inter.start,row.df$inter.end))
      ovp.df <- findOverlapPairs(promint.gr,cvar.gr) %>% as.data.frame(.)
      ovp.df <- ovp.df[!duplicated(ovp.df),] # promoter interactions overlapping SNP/enh bait 
      enhint.gr <- GRanges(seqnames=ovp.df$second.seqnames,
                           IRanges(ovp.df$second.start,ovp.df$second.end))
      ovp.df2 <- findOverlapPairs(enhint.gr,prom.gr) %>% as.data.frame(.)
      ovp.df2 <- ovp.df2[!duplicated(ovp.df2),] # SNP/enh interaction overlapping with promoter  
      if (dim(ovp.df2)[1]>0){
        print(cap)
      }
    }
  }
}


```



