---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup 


```{r}

"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")
library("ggbio")
library("RColorBrewer")
library("viridis")
library("Homo.sapiens")
library("GenomicRanges")

serv.dir1 <- "/home/jason/science/servers/FUSE5/"
serv.dir2 <- "/home/jason/science/servers/FUSE2/"
got2d.dir <- "/home/jason/science/servers/FUSE/" 
cbrg.dir.prom <- serv.dir2 %&% "wtsa/promoter_driven/" 
cbrg.dir.enh <- serv.dir2 %&% "wtsa/enhancer_driven/first_round/" 
rescomp.dir.prom <- serv.dir1 %&%
  "projects/wtsa/promoter_driven/statistical_analysis/output_files/"
rescomp.dir.enh <- serv.dir1 %&%
  "projects/wtsa/enhancer_driven/first_round/statistical_anlayses/output_files/"
param.file.prom <- cbrg.dir.prom %&% "capture_compare_parameters.txt"
param.file.enh <- cbrg.dir.enh %&% "capture_compare_parameters_pruned_chr.txt"

gencred.df <- fread(serv.dir1 %&% "datasets/diamante_hrc/gencred.txt")

ref.eqtl.df <- cbrg.dir.prom %&%
  "reference_files/eQTL-index.bed" %>% fread(.)
ref.gwas.df <- cbrg.dir.prom %&%
  "reference_files/gwas-index.bed" %>% fread(.)

data(genesymbol, package = "biovizBase")

plot.dir <-  serv.dir1 %&%
  "projects/wtsa/enhancer_driven/first_round/statistical_anlayses/plots/"

```


```{r}


science_theme <- theme(
  panel.grid.major = element_line(size = 0.5, color = "grey"),
  panel.grid.minor = element_blank(),
  plot.title = element_text(hjust = 0.5),
  text = element_text(size = 14),  axis.line = element_line(color="black", size = 0.7),
  axis.line.x = element_line(color="black", size = 0.7),
  axis.line.y = element_line(color="black", size = 0.7),
  # plot.margin = unit(c(0.7,0.7,0.7,0.7), "lines"),
  panel.border=element_blank(),
  strip.background = element_blank()
)

```


# Plotting Functions 

```{r}

prepare_plot_list <- function(id,experiment="promoter"){
  # Experiment can either be "promoter" or "enhancer"
  if (experiment=="promoter"){
    winfile.dir <- cbrg.dir.prom %&% "PromCap_cis_analysis/4_plotting/C_windowed/"
    parameters.file <- param.file.prom
  } else if (experiment=="enhancer"){
    winfile.dir <- cbrg.dir.enh %&% "capture_compare_cis_analysis/4_plotting/C_windowed/"
    parameters.file <- param.file.enh
  } else{
    print("Experiment can either be 'promoter' or 'enhancer'")
  }
  data.file <- winfile.dir %&% id %&% "_window.tab"
  data <- as.tibble(read.table(data.file, header=T))
  parameters <- as.tibble(read.table(parameters.file))
  names(parameters) <- c("Viewpoint", "Chr", "Frag_start", "Frag_stop", "Exclusion_Start",
                         "Exclusion_Stop", "Plot_Region_Start",
                         "Plot_Region_Stop", "Bin_Size", "Window_size")  
  # select viewpoint from parameters file
  viewp <- parameters %>% filter(Viewpoint == id)
  # select exclusion fragments within plotting region
  parameters <- parameters %>%
  filter(Frag_stop >= viewp$Plot_Region_Start & Frag_start <= viewp$Plot_Region_Stop)
  # add window size to exclusion zone only for viewpoint
  parameters <- parameters %>%
    mutate(Exclusion_Start = if_else(Viewpoint == id, (Exclusion_Start - Window_size),
                                     Exclusion_Start)) %>%
  mutate(Exclusion_Stop = if_else(Viewpoint == id, (Exclusion_Stop + Window_size),
                                  Exclusion_Stop))
  to.exclude <- parameters %>% filter(Viewpoint == id)
  # Gather Data =================================================================
  # 1 gather in long format
  d <- data %>% gather(key, value, -c(BinNum, Chr, Start, Stop))
  # 2 extract condition and replicate
  d <- d %>% 
    mutate(condition = sub('_.+', "", key, perl=T)) %>%
    mutate(replicate = sub('.+_', "", key, perl=T)) %>%
    mutate(pos = Start + (Stop - Start)/2)
  # Calculate Mean and STDEV ====================================================
  d <- d %>%
    group_by(BinNum, Chr, Start, Stop, pos, condition) %>%
  summarize(mean = mean(value), sd = sd(value)) %>%
  ungroup()  
  return(list(d,to.exclude,parameters))
}

```

```{r}

capc_plot <- function(id,experiment,plotwin=5e5,plotmax=NULL,endoOnly=FALSE,
                      promrefsnps=FALSE){
  plt.list <- prepare_plot_list(id,experiment)
  plt.list[[3]] <- plt.list[[3]][grepl(id,plt.list[[3]]$Viewpoint),]
  vp = plt.list[[3]]$Frag_start
  d <- plt.list[[1]]; to.exclude <- plt.list[[2]]; 
  params <- plt.list[[3]]
  if (experiment=="promoter"){
    d$condition <- factor(d$condition,levels=c("EndoB","Blymph"))
    bg.dir <- cbrg.dir.prom %&%
      "PromCap_cis_analysis/2_unionBedgraphs/B_normalised_counts/"
    peakc.mod.df <- rescomp.dir.prom %&% 
      "peakC-modeled-default_EndoC.txt" %>% fread(.)
    peakc.comp.df <- rescomp.dir.prom %&% 
      "peakC-comparative-default.txt" %>% fread(.)
  } else if(experiment=="enhancer"){
    d$condition <- factor(d$condition,levels=c("Endo","hESC"))
    bg.dir <- cbrg.dir.enh %&%
      "capture_compare_cis_analysis/2_unionBedgraphs/B_normalised_counts/"
    peakc.mod.df <- rescomp.dir.enh %&% 
      "peakC-modeled-default_EndoC.txt" %>% fread(.)
    peakc.comp.df <- rescomp.dir.enh %&% 
      "peakC-comparative-default.txt" %>% fread(.)
  } else{
    print("Experiment can either be 'promoter' or 'enhancer'")
  }
  
  if (endoOnly==TRUE){
    d <- filter(d,grepl("Endo",condition))
  }

  bdg.df <- fread(bg.dir %&% id %&% "_normalised.unionbdg") %>% as.data.frame(.)
  
  if (is.null(plotmax)==TRUE){
    plotmax <- filter(d,Stop<to.exclude$Exclusion_Start |
                      Start>to.exclude$Exclusion_Stop)$mean %>% 
                      max(.) %>% mean(.) %>% round(.)
    plotmax <- plotmax+50    
  }


  p <- ggplot(filter(d,Stop<to.exclude$Exclusion_Start |
                      Start>to.exclude$Exclusion_Stop),
              aes(x=pos, y=mean, col = condition, fill = condition)) +
    geom_ribbon(inherit.aes = F, 
                aes(x=pos, ymin=mean-sd, ymax=mean+sd, fill=condition), alpha=.35) + 
    geom_line() + 
    scale_fill_manual(values=viridis(20)[c(3,9)],name="Cell Type") +
    scale_color_manual(values=viridis(20)[c(3,9)],name="Cell Type") +
    geom_rect(data=to.exclude, inherit.aes=F, 
              aes(xmin=Exclusion_Start, xmax=Exclusion_Stop), 
              ymin=-25, ymax=max(d$mean)+max(d$sd), col="lightgrey", fill="white") + 
    geom_rect(data=params, inherit.aes=F, 
              aes(xmin=Frag_start, xmax=Frag_stop), ymin=-25, 
              ymax=max(d$mean)+max(d$sd), col="grey", fill="grey") + 
    labs(x="Genomic Position", y="Mean Interaction") + 
    ggtitle(id) +
    theme_bw() + science_theme +  
    theme(panel.grid = element_blank())	 
  
  peakc.mod <- peakc.mod.df$peak.start
  peakc.comp <- peakc.comp.df$peak.start 
  peakc.df <- filter(bdg.df,V2 %in% peakc.mod) %>% 
    dplyr::select(.,one_of("V1","V2","V3")) %>% 
    filter(.,V3<to.exclude$Exclusion_Start |
                      V3>to.exclude$Exclusion_Stop)
  comp.df <- filter(bdg.df,V2 %in% peakc.comp) %>% 
    dplyr::select(.,one_of("V1","V2","V3")) %>% 
    filter(.,V3<to.exclude$Exclusion_Start |
                      V3>to.exclude$Exclusion_Stop)
  if (dim(peakc.df)[1]>0){
    p = p + geom_rect(data=peakc.df,inherit.aes=F,aes(xmin=V2,xmax=V3),
                ymin=-(plotmax/50),ymax=(plotmax/50),size=0.1,
                color="firebrick3",fill="firebrick1")
  }
  if (dim(comp.df)[1]>0){
    p = p + geom_rect(data=comp.df,inherit.aes=F,aes(xmin=V2,xmax=V3),
                ymin=-(plotmax/50),ymax=(plotmax/50),size=0.1,
                color=viridis(20)[20],fill=viridis(20)[15])    
  }
  p <- p + coord_cartesian(xlim=c(params$Frag_start-(plotwin), 
                           params$Frag_stop+(plotwin)),
                  ylim=c(0,plotmax))
  return(p)
}

```

```{r}

gene_plot <- function(id,experiment,plotwin=5e5){
  plt.list <- prepare_plot_list(id,experiment)
  plt.list[[3]] <- plt.list[[3]][grepl(id,plt.list[[3]]$Viewpoint),]
  params <- plt.list[[3]]
  reg.gr <- GRanges(seqnames=params$Chr,
                 IRanges(params$Frag_start-(plotwin/2),
                 params$Frag_stop+(plotwin/2)))
  wh <- genesymbol[genesymbol %over% reg.gr] 
  wh <- range(wh, ignore.strand = TRUE)
  p <- autoplot(Homo.sapiens, which = wh,
                color=cividis(20)[1],
                fill=cividis(20)[4]) 
  p = p + theme_bw() + 
    theme(panel.grid=element_blank()) +
    ylab("Transcripts") + 
    xlab("Genomic Position")
  return(p)
}

```

```{r}

credset_plot <- function(id,experiment,plotwin=5e5,
                         promrefsnps=FALSE){
  plt.list <- prepare_plot_list(id,experiment)
  plt.list[[3]] <- plt.list[[3]][grepl(id,plt.list[[3]]$Viewpoint),]
  params <- plt.list[[3]]
  sub <- filter(gencred.df,
                CHR==params$Chr,
                POS>=params$Frag_start-(plotwin),
                POS<=params$Frag_stop+(plotwin))
  plt <- ggplot(data=sub,aes(x=POS,y=PPA)) + 
    geom_point(shape=21,color=inferno(20)[2],
               fill=inferno(20)[8],size=1.5) + 
    scale_y_continuous(breaks=seq(0,1,0.2),limits=c(0,1.15)) +
    theme(panel.grid.major = element_blank())
  if (promrefsnps==TRUE){
    s1 <- filter(ref.eqtl.df,V1==params$Chr,
                 V2>=params$Frag_start-(plotwin),
                 V2<=params$Frag_stop+(plotwin))
    s2 <- filter(ref.gwas.df,V1==params$Chr,
                 V2>=params$Frag_start-(plotwin),
                 V2<=params$Frag_stop+(plotwin))
    if (dim(s1)[1]>0){
      plt <- plt + 
        geom_segment(data=s1,
                     aes(x=V2,xend=V2),y=1.15,yend=1.05,
                     arrow=arrow(),color=viridis(20)[6])
    }
    if (dim(s2)[1]>0){
      plt <- plt + 
        geom_segment(data=s2,
                     aes(x=V2,xend=V2),y=1.12,yend=1.02,
                     arrow=arrow(),color=viridis(20)[15])
    }
  }
  plt = plt + theme_bw()
  return(plt)
}

```


 Testing plotting functions, looks good 

```{r}

plt1 <- capc_plot(id="ADCY5",experiment="promoter",plotwin=5e5,endoOnly = TRUE)
plt2 <- capc_plot(id="ADCY5__1TG__te__chr3_123054770",
                  experiment="enhancer",plotwin=5e5,endoOnly = TRUE)
cplt <- credset_plot(id="ADCY5",experiment="promoter",plotwin=5e5,promrefsnps=TRUE)
gplt <- gene_plot(id="ADCY5",experiment="promoter",plotwin=5e5)
plt <- tracks(cplt,plt1,plt2,gplt,heights=c(0.8,1,1,1.5)) + 
  scale_x_sequnit(unit="Mb") 


```


# Promoter analysis 

21 eQTL/GWAS colocalized signal eGene promoters captured 

```{r}

check_for_snpoverlap <- function(par.df,peakc.df,gene.vec,win=0){
  param.sub <- filter(par.df,V1 %in% gene.vec) # 
  peakc.sub <- filter(peakc.df,bait %in% gene.vec)
  out.df <- c()
  for (i in 1:dim(peakc.sub)[1]){
    row.df <- peakc.sub[i,]
    eqtls <- filter(ref.eqtl.df,V1==row.df$chrom,
                    V2>=(row.df$peak.start-win),V2<=(row.df$peak.end+win)) %>% 
      dplyr::select(.,one_of("V2","V4"))
    names(eqtls) <- c("eqtl.pos","eqtl.name")
    if (dim(eqtls)[1]>0){
      build.df <- cbind(row.df,eqtls)
      out.df <- rbind(out.df,build.df)
    }
  }
  return(out.df)
}

```


```{r}

param.df1 <- fread(param.file.prom)
param.df2 <- fread(param.file.enh)

peakc.mod.df <- rescomp.dir.prom %&% 
  "peakC-modeled-default_EndoC.txt" %>% fread(.)
peakc.comp.df <- rescomp.dir.prom %&% 
  "peakC-comparative-default.txt" %>% fread(.)
# Subset of eGene promoters that overlap MVD Bunt 2015 Plot Genet, 
# Supp Table 2, 21 colocalized eQTLS
mvdb.egenes <- c("ABO","NKX6-3","AP3S2","STARD10","CAMK1D", #"C15orf38-ap3s2",
                 "DGKB","ACP2","MADD","WARS","ADCY5","RBM6","FADS1",
                 "MTNR1B","CTD-2260A17.2","MGAT5","LINC00927","ZMIZ") # 

direct.overlap <- check_for_snpoverlap(param.df1,
                                       peakc.mod.df,mvdb.egenes)
direct.overlap.comp <- check_for_snpoverlap(param.df1,
                                       peakc.comp.df,mvdb.egenes)
oneKb.overlap <- check_for_snpoverlap(param.df1,peakc.mod.df,
                                      mvdb.egenes,win=1000)

```

NOTE: NKX6-3 Promoter physically interacts with eQTL from comparative interaction 


```{r}

plt1a <- capc_plot(id="ADCY5",experiment="promoter",plotwin=2.1e5,endoOnly = TRUE)
plt2a <- capc_plot(id="CAMK1D",experiment="promoter",plotwin=2.5e5,endoOnly = TRUE)
plt3a <- capc_plot(id="STARD10",experiment="promoter",plotwin=2.5e5,endoOnly = TRUE)

plt1b <- credset_plot(id="ADCY5",experiment="promoter",
                     plotwin=2.1e5,promrefsnps=TRUE)
plt2b <- credset_plot(id="CAMK1D",experiment="promoter",
                     plotwin=2.5e5,promrefsnps=TRUE)
plt3b <- credset_plot(id="STARD10",experiment="promoter",
                     plotwin=2.5e5,promrefsnps=TRUE)

plt1c <- gene_plot(id="ADCY5",experiment="promoter",plotwin=2.1e5)
plt2c <- gene_plot(id="CAMK1D",experiment="promoter",plotwin=2.5e5)
plt3c <- gene_plot(id="STARD10",experiment="promoter",plotwin=2.5e5)

p1 <- tracks(plt1b,plt1a,plt1c,heights=c(0.2,1,0.5)) +
  scale_x_sequnit(unit="Mb") 
ggsave(plot=p1,filename = plot.dir %&% "tracks_prom_ADCY5.png",
       height=10,width=10)
ggsave(plot=p1,filename = plot.dir %&% "tracks_prom_ADCY5.pdf",
       height=10,width=10)

p2 <- tracks(plt2b,plt2a,plt2c,heights=c(0.2,1,0.5)) +
  scale_x_sequnit(unit="Mb") 
ggsave(plot=p2,filename = plot.dir %&% "tracks_prom_CAMK1D.png",
       height=10,width=10)
ggsave(plot=p2,filename = plot.dir %&% "tracks_prom_CAMK1D.pdf",
       height=10,width=10)

p3 <- tracks(plt3b,plt3a,plt3c,heights=c(0.2,1,0.5)) +
  scale_x_sequnit(unit="Mb") 
ggsave(plot=p3,filename = plot.dir %&% "tracks_prom_STARD10.png",
       height=10,width=10)
ggsave(plot=p3,filename = plot.dir %&% "tracks_prom_STARD10.pdf",
       height=10,width=10)

# Just checking NKX6-3 out of curiosity 
plt4a <- capc_plot(id="NKX6-3",experiment="promoter",plotwin=2.5e5)
plt4b <- credset_plot(id="NKX6-3",experiment="promoter",
                     plotwin=2.5e5,promrefsnps=TRUE)
plt4c <- gene_plot(id="NKX6-3",experiment="promoter",plotwin=2.5e5)
p4 <- tracks(plt4b,plt4a,plt4c,heights=c(0.2,1,0.5)) +
  scale_x_sequnit(unit="Mb") 
ggsave(plot=p4,filename = plot.dir %&% "tracks_prom_NKX6-3.png",
       height=10,width=10)
ggsave(plot=p4,filename = plot.dir %&% "tracks_prom_NKX6-3.pdf",
       height=10,width=10)

# Just checking KCNK17 out of curiosity 
plt8a <- capc_plot(id="DNLZ",experiment="promoter",plotwin=6e4)
plt8b <- credset_plot(id="DNLZ",experiment="promoter",
                     plotwin=6e4,promrefsnps=TRUE)
plt8c <- gene_plot(id="DNLZ",experiment="promoter",plotwin=6e4)
p8 <- tracks(plt8b,plt8a,plt8c,heights=c(0.2,1,0.5)) +
  scale_x_sequnit(unit="Mb") 
ggsave(plot=p8,filename = plot.dir %&% "tracks_prom_DNLZ.png",
       height=10,width=10)
ggsave(plot=p8,filename = plot.dir %&% "tracks_prom_DNLZ.pdf",
       height=10,width=10)

```

Look up distance of interactions to bait 

```{r}

ref.eqtl.df$distToBait <- map(1:dim(ref.eqtl.df)[1],function(i){
  row.df <- ref.eqtl.df[i,]
  egene <- strsplit(row.df$V4,split="_")[[1]][2]
  egene <- ifelse(egene=="C15orf38-AP3S2","C15orf38-ap3s2",egene)
  sub <- filter(param.df1,V1==egene)
  gr1 <- GRanges(seqnames=row.df$V1,
                 IRanges(row.df$V2,row.df$V3))
  gr2 <- GRanges(seqnames=sub$V2,
                 IRanges(sub$V3,sub$V4))
  dist <- distance(gr1,gr2)
}) %>% as.integer(.)

ref.eqtl.df$egene <- map(1:dim(ref.eqtl.df)[1],function(i){
  row.df <- ref.eqtl.df[i,]
  egene <- strsplit(row.df$V4,split="_")[[1]][2]
}) %>% as.character(.)

```


```{r}

ref.eqtl.df$egene %>% length(.)
ref.eqtl.df$egene %>% unique(.) %>% length(.)
# Captured 21 eGenes corresponding to 27 eQTLs 
egene.vec <- ref.eqtl.df$egene %>% unique(.)

direct.overlap2 <- check_for_snpoverlap(param.df1,peakc.mod.df,
                                       egene.vec)
oneKb.overlap2 <- check_for_snpoverlap(param.df1,peakc.mod.df,
                                       egene.vec,win=1000)
direct.comp.overlap2 <- check_for_snpoverlap(param.df1,peakc.comp.df,
                                       egene.vec)
oneKb.comp.overlap2 <- check_for_snpoverlap(param.df1,peakc.comp.df,
                                       egene.vec,win=1000)


summary(ref.eqtl.df$distToBait) # med dist 30Kb, 3rd Qu. 76.5Kb

# 72-101Kb for three that sig overlap 
# Larp about 30Kb away 

```

#1    DNLZ  chr9  139233724 139257349 23626 139248082 rs28642213_GPSM1
#2    DNLZ  chr9  139233724 139257349 23626 139246594  rs28679497_DNLZ
#3    DNLZ  chr9  139233724 139257349 23626 139246768 rs61386106_CARD9
#37 KCNK17  chr6   39283118  39292854  9737  39290834 rs3807042_KCNK17
#48 NKX6-3  chr8   41505686  41512695  7010  41509915 rs4736819_NKX6-3


```{r}

plt5a <- capc_plot(id="LARP6",experiment="promoter",plotwin=2.5e5)
plt5b <- credset_plot(id="LARP6",experiment="promoter",
                     plotwin=2.5e5,promrefsnps=TRUE)
plt5c <- gene_plot(id="LARP6",experiment="promoter",plotwin=2.5e5)
#p5 <- tracks(plt5a,plt5c,heights=c(1,1)) +
#  scale_x_sequnit(unit="Mb") 

```


# Now see if sig interactions reciprocated from enhancer capture 

```{r}

# find enhancer baits that overlap eQTLs from promoter capture interactions

find_out_enh_bait <- function(dirover,param.df){
  out.df <- c()
  for (i in 1:dim(direct.overlap2)[1]){
    row.df <- dirover[i,]
    sub <- filter(param.df2,V2==row.df$chrom,V3<=row.df$eqtl.pos,V4>=row.df$eqtl.pos) %>% 
      dplyr::select(.,one_of("V1","V3","V4"))
    names(sub) <- c("enh.bait","enh.bait.start","enh.bait.end")
    if (dim(sub)[1]>0){
      build.df <- cbind(row.df,sub)
      out.df <- rbind(out.df,build.df)
    }
  } 
  return(out.df)
}

get_recip_inter <- function(enh.cap.df,enh.peakc.df,prom.param.df){
  out.df <- c()
  for (i in 1:dim(enh.cap.df)[1]){
    row.df <- enh.cap.df[i,]
    prom.bait.start <- filter(prom.param.df,V1==row.df$bait)$V3
    prom.bait.end <- filter(prom.param.df,V1==row.df$bait)$V4
    prom.bait.gr <- GRanges(seqnames=row.df$chrom,
                            IRanges(prom.bait.start,prom.bait.end))
    sub <- filter(enh.peakc.df,bait==row.df$enh.bait)
    inter.gr <- GRanges(seqnames = sub$chrom,
                        IRanges(sub$peak.start,sub$peak.end))
    over.df <- findOverlapPairs(prom.bait.gr,inter.gr) %>% as.data.frame(.) %>% 
      dplyr::select(.,one_of("second.start","second.end"))
    names(over.df) <- c("enh.peak.start","enh.peak.end")
    if (dim(over.df)[1]>0){
      build.df <- cbind(row.df,over.df)
      out.df <- rbind(out.df,build.df)
    }
  }
  return(out.df)
}


```


```{r}



enh.cap.df <- find_out_enh_bait(direct.overlap2,fread(param.file.enh))
enh.peakc.df <- rescomp.dir.enh %&% 
    "peakC-modeled-default_EndoC.txt" %>% fread(.)
prom.param.df <- param.df1
recip.df <- get_recip_inter(enh.cap.df,enh.peakc.df,prom.param.df) # 

```


```{r}

#enh.peakc.df2 <- rescomp.dir.enh %&% 
#    "peakC-modeled-default_hESC.txt" %>% fread(.)
#get_recip_inter(enh.cap.df,enh.peakc.df2,prom.param.df) # 
# Save when use hESC instead of EndoC

```


```{r}

plt6a <- credset_plot(id="ADCY5",experiment="promoter",
                     plotwin=2.5e5,promrefsnps=TRUE)
plt6b <- capc_plot(id="ADCY5",experiment="promoter",plotwin=2.5e5,endoOnly=TRUE)
plt6c <- capc_plot(id="ADCY5__24DF__cv__chr3_123065778",experiment="enhancer",plotwin=2.5e5,endoOnly=TRUE)
plt6d <- gene_plot(id="ADCY5",experiment="promoter",plotwin=2.5e5)
p6 <- tracks(plt6a,plt6b,plt6c,plt6d,heights=c(0.2,1,1,0.5)) +
  scale_x_sequnit(unit="Mb") 
ggsave(plot=p6,filename = plot.dir %&% "tracks_PromEnh_ADCY5.png",
       height=10,width=20)
ggsave(plot=p6,filename = plot.dir %&% "tracks_PromEnh_ADCY5.pdf",
       height=10,width=20)
```


```{r}

plt7a <- credset_plot(id="CAMK1D",experiment="promoter",
                     plotwin=2.5e5,promrefsnps=TRUE)
plt7b <- capc_plot(id="CAMK1D",experiment="promoter",plotwin=2.5e5,endoOnly=TRUE)
plt7c <- capc_plot(id="rs11257658_CAMK1D__5QTL__eQTL__chr10_12309268",experiment="enhancer",plotwin=2.5e5,endoOnly=TRUE)
plt7d <- gene_plot(id="CAMK1D",experiment="promoter",plotwin=2.5e5)
p7 <- tracks(plt7a,plt7b,plt7c,plt7d,heights=c(0.2,1,1,0.5)) +
  scale_x_sequnit(unit="Mb") 
ggsave(plot=p7,filename = plot.dir %&% "tracks_PromEnh_CAMK1D.png",
       height=10,width=15)
ggsave(plot=p7,filename = plot.dir %&% "tracks_PromEnh_CAMK1D.pdf",
       height=10,width=15)
# NOTE CAMK1D was poorly captured in experiment 

```

```{r}

plt8a <- credset_plot(id="STARD10",experiment="promoter",
                     plotwin=2.5e5,promrefsnps=TRUE)
plt8b <- capc_plot(id="STARD10",experiment="promoter",plotwin=2.5e5,endoOnly=TRUE)
plt8c <- capc_plot(id="ARAP1__99DF__cv__chr11_72432985",experiment="enhancer",plotwin=2.5e5,endoOnly=TRUE)
plt8d <- gene_plot(id="STARD10",experiment="promoter",plotwin=2.5e5)
p8 <- tracks(plt8a,plt8b,plt8c,plt8d,heights=c(0.2,1,1,0.5)) +
  scale_x_sequnit(unit="Mb") 
ggsave(plot=p8,filename = plot.dir %&% "tracks_PromEnh_STARD10.png",
       height=10,width=15)
ggsave(plot=p8,filename = plot.dir %&% "tracks_PromEnh_STARD10.pdf",
       height=10,width=15)

# Can make a case for STARD10 being replicated, just not quite with peakC

```



```{r}

plt9a <- credset_plot(id="FBRSL1__109DF__cv__chr12_133069698",experiment="enhancer",
                     plotwin=7e5,promrefsnps=TRUE)
plt9b <- capc_plot(id="FBRSL1__109DF__cv__chr12_133069698",experiment="enhancer",plotwin=7e5,endoOnly=TRUE)
plt9c <- capc_plot(id="FBRSL1__109DF__cv__chr12_133066392",experiment="enhancer",plotwin=7e5,endoOnly=TRUE)
plt9d <- gene_plot(id="FBRSL1__109DF__cv__chr12_133069698",experiment="enhancer",plotwin=7e5)
p9 <- tracks(plt9a,plt9b,plt9c,plt9d,heights=c(0.2,1,1,0.5)) +
  scale_x_sequnit(unit="Mb") 
ggsave(plot=p9,filename = plot.dir %&% "tracks_CVEnh_FBRSL1.png",
       height=10,width=15)
ggsave(plot=p9,filename = plot.dir %&% "tracks_CVEnh_FBRSL1.pdf",
       height=10,width=15)


```

