---
title: "update-enhancer-oligo-files"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup 

```{r}
"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")
library("GenomicRanges")

local.dir <- "/Users/jasont/science/projects/wtsa/"
rescomp.dir <- "/Users/jasont/science/servers/FUSE5/"
cbrg.dir <- "/Users/jasont/science/servers/FUSE2/"


```


```{r}
e1.dir <- cbrg.dir %&% "wtsa/enhancer_driven/first_round/"
e1.cc.file  <- e1.dir %&% "Oligofile-original.txt"
e1.df <- fread(e1.cc.file) 
e2.df <- fread(rescomp.dir %&% "projects/wtsa/enhancer_driven/" %&%
              "second_round/rerun_pipeline/OligoFileRedo.txt")
```


## Determine what column manipulations need to be made 

```{r}

all(e1.df$V6 == (e1.df$V3 - 1000))  
all(e1.df$V7 == (e1.df$V4 + 1000))

all(e2.df$V6 == (e2.df$V3 - 1000)) # Need to update fields V5 and V6 for enhancer capture second round
all(e2.df$V7 == (e2.df$V4 + 1000))

```

```{r}
e1.update.df <- e1.df
### e1.update.df$V6 <- (e1.df$V3 - 1000) # unnecessary 
### e1.update.df$V7 <-  (e1.df$V4 + 1000)

e2.update.df <- e2.df
e2.update.df$V6 <- (e2.df$V3 - 1000)
e2.update.df$V7 <-  (e2.df$V4 + 1000)

all(e1.update.df$V6 == (e1.update.df$V3 - 1000)) 
all(e1.update.df$V7 == (e1.update.df$V4 + 1000))

all(e2.update.df$V6 == (e2.update.df$V3 - 1000)) 
all(e2.update.df$V7 == (e2.update.df$V4 + 1000))
```

Both E1 and E2 data frames have been evaluated and confirmed that they 
now include a 1kb exclusion zone

## Determine what captures need to be merged 

```{r}
e1.cap.gr <- GRanges(seqnames=e1.update.df$V2,
                 IRanges(start=e1.update.df$V3,end=e1.update.df$V4))
length(e1.cap.gr) == length(reduce(e1.cap.gr)) # FALSE 
e1.excl.gr <- GRanges(seqnames=e1.update.df$V5,
                 IRanges(start=e1.update.df$V6,end=e1.update.df$V7))
length(e1.excl.gr) == length(reduce(e1.excl.gr)) # FALSE; 238 vs 149
```

There captures that overlap with other capture exclusion zones, in the first round set 

```{r}
e2.cap.gr <- GRanges(seqnames=e2.update.df$V2,
                 IRanges(start=e2.update.df$V3,end=e2.update.df$V4))
length(e2.cap.gr) == length(reduce(e2.cap.gr)) # FALSE; 204 vs 203 
e2.excl.gr <- GRanges(seqnames=e2.update.df$V5,
                 IRanges(start=e2.update.df$V6,end=e2.update.df$V7))
length(e2.excl.gr) == length(reduce(e2.excl.gr)) # FALSE; 204 vs 191 
```

There are captures that overlap with other capture exclusion zones, in the second round set 

Now, let's determine what needs to be merged, will iterate through the reduced 
exclusion set, look up overlapping captures, and create a new merged ID, and write all output 

### First Round Enhancer 

```{r}
e1.output.df <- c()
e1.merged.ref.df <- c()
names(e1.excl.gr) <- e1.update.df$V1
e1.excl.reduce <- reduce(e1.excl.gr)
pb <- txtProgressBar(min=0,max=length(e1.excl.reduce),style=3)
for (i in 1:length(e1.excl.reduce)){
  setTxtProgressBar(pb,i)
  query.gr <- e1.excl.reduce[i]
  sub.gr <- e1.excl.gr[e1.excl.gr %over% query.gr]
  if (length(sub.gr)==1){
    cap.id <- names(sub.gr)
    build.df <- filter(e1.update.df,V1==cap.id)
    build.df$V3 <- as.integer(build.df$V3)
    build.df$V4 <- as.integer(build.df$V4)
    build.df$V2 <- as.character(build.df$V2)
    build.df$V5 <- as.character(build.df$V5)
    e1.output.df <- rbind(e1.output.df,build.df)
  } else{
    cap.id.vec <- names(sub.gr)
    query.df <- query.gr %>% as.data.frame(.)
    chrom <- query.df$seqnames
    excl.start <- query.df$start; excl.end <- query.df$end
    cap.start <- excl.start + 1000; cap.end <- excl.end - 1000
    ### win.start <- cap.start - 1e6; win.end <- cap.end + 1e6 
    cap.id <- "merged_E1_chr"%&%chrom%&%"_"%&%cap.start%&%"-"%&%cap.end
    build.df <- data.frame(V1=cap.id,V2=chrom,V3=cap.start,V4=cap.end,
                           V5=chrom,V6=excl.start,V7=excl.end,
                           V8=1,V9="A",stringsAsFactors = F)
    build.df$V3 <- as.integer(build.df$V3)
    build.df$V4 <- as.integer(build.df$V4)
    build.df$V2 <- as.character(build.df$V2)
    build.df$V5 <- as.character(build.df$V5)
    e1.output.df <- rbind(e1.output.df,build.df)
    build.df2 <- data.frame("capture.ID"=cap.id.vec,"merged.ID"=cap.id,
                            stringsAsFactors = F)
    e1.merged.ref.df <- rbind(e1.merged.ref.df,build.df2)
  }
}
```


```{r}
all(e1.output.df$V6 == (e1.output.df$V3 - 1000)) 
all(e1.output.df$V7 == (e1.output.df$V4 + 1000))
```


```{r}
write.table(e1.output.df,file=rescomp.dir %&% "projects/wtsa/capture_c_rerun/input_files/" %&%
              "captureC_E1_oligofile.txt",sep="\t",quote=F,row.names=F,col.names=F)
write.table(e1.merged.ref.df,file=rescomp.dir %&% "projects/wtsa/capture_c_rerun/input_files/" %&%
              "captureC_E1_merged-keyfile.txt",sep="\t",quote=F,row.names=F,col.names=F)
```

### Second Round Enhancer 

```{r}
e2.output.df <- c()
e2.merged.ref.df <- c()
names(e2.excl.gr) <- e2.update.df$V1
e2.excl.reduce <- reduce(e2.excl.gr)
pb <- txtProgressBar(min=0,max=length(e2.excl.reduce),style=3)
for (i in 1:length(e2.excl.reduce)){
  setTxtProgressBar(pb,i)
  query.gr <- e2.excl.reduce[i]
  sub.gr <- e2.excl.gr[e2.excl.gr %over% query.gr]
  if (length(sub.gr)==1){
    cap.id <- names(sub.gr)
    build.df <- filter(e2.update.df,V1==cap.id)
    build.df$V3 <- as.integer(build.df$V3)
    build.df$V4 <- as.integer(build.df$V4)
    build.df$V2 <- as.character(build.df$V2)
    build.df$V5 <- as.character(build.df$V5)
    e2.output.df <- rbind(e2.output.df,build.df)
  } else{
    cap.id.vec <- names(sub.gr)
    query.df <- query.gr %>% as.data.frame(.)
    chrom <- query.df$seqnames
    excl.start <- query.df$start; excl.end <- query.df$end
    cap.start <- excl.start + 1000; cap.end <- excl.end - 1000
    ### win.start <- cap.start - 1e6; win.end <- cap.end + 1e6 
    cap.id <- "merged_E2_chr"%&%chrom%&%"_"%&%cap.start%&%"-"%&%cap.end
    build.df <- data.frame(V1=cap.id,V2=chrom,V3=cap.start,V4=cap.end,
                           V5=chrom,V6=excl.start,V7=excl.end,
                           V8=1,V9="A",stringsAsFactors = F)
    build.df$V3 <- as.integer(build.df$V3)
    build.df$V4 <- as.integer(build.df$V4)
    build.df$V2 <- as.character(build.df$V2)
    build.df$V5 <- as.character(build.df$V5)
    e2.output.df <- rbind(e2.output.df,build.df)
    build.df2 <- data.frame("capture.ID"=cap.id.vec,"merged.ID"=cap.id,
                            stringsAsFactors = F)
    e2.merged.ref.df <- rbind(e2.merged.ref.df,build.df2)
  }
}
```


```{r}
all(e2.output.df$V6 == (e2.output.df$V3 - 1000)) 
all(e2.output.df$V7 == (e2.output.df$V4 + 1000))
```


```{r}
write.table(e2.output.df,file=rescomp.dir %&% "projects/wtsa/capture_c_rerun/input_files/" %&%
              "captureC_E2_oligofile.txt",sep="\t",quote=F,row.names=F,col.names=F)
write.table(e2.merged.ref.df,file=rescomp.dir %&% "projects/wtsa/capture_c_rerun/input_files/" %&%
              "captureC_E2_merged-keyfile.txt",sep="\t",quote=F,row.names=F,col.names=F)
```


### Evaluate capture fragment distribution lengths

```{r}
# not merged 
dist1.nm <- filter(e1.output.df,!grepl("merged",V1))$V4-filter(e1.output.df,!grepl("merged",V1))$V3
summary(dist1.nm)
## Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##   83.0   479.5   865.0   889.6  1255.0  1964.0 
# merged 
dist1.m <- filter(e1.output.df,grepl("merged",V1))$V4-filter(e1.output.df,grepl("merged",V1))$V3
summary(dist1.m)
## Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##  866    2401    3200    3849    4564    8574 

```

```{r}
# not merged 
dist2.nm <- filter(e2.output.df,!grepl("merged",V1))$V4-filter(e2.output.df,!grepl("merged",V1))$V3
summary(dist2.nm)
## Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##  149.0   459.5   773.5   937.9  1283.0  3110.0 
# merged 
dist2.m <- filter(e2.output.df,grepl("merged",V1))$V4-filter(e2.output.df,grepl("merged",V1))$V3
summary(dist2.m)
## Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##    822    1501    2070    2154    2654    3640 

```

In total, there are 149 capture regions to process in the first round and 
there are 191 capture regions to process in the second round

