---
title: "Make Capsequm Bed file"
author: "Jason Torres"
date: "June 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
"%&%" <- function(a,b) paste0(a,b) 
library("data.table")
library("dplyr")

serv.dir <- "/Users/jtorres/FUSE/"
rds.dir <- work.dir %&% "RDS/"
txt.dir <- work.dir %&% "txt/"

```


# First Round Capture variants BED File 


```{r}

df <- fread(txt.dir%&%"first-round-variants.txt")
CHROM <- "chr" %&%  df$CHR
POS0 <- df$POS - 1
POS1 <- df$POS
NAME <- df$Locus %&% "," %&% df$SEGNUMBER %&% "," %&% df$Type %&% "," %&% CHROM %&% ":" %&% POS1 
bed.df <- data.frame(CHROM,POS0,POS1,NAME,stringsAsFactors = FALSE)
write.table(x=bed.df,file = txt.dir %&% "first-round-variants.bed",
            sep="\t",quote=FALSE,col.names=FALSE,row.names=FALSE)

```



# File to share 

```{r}

additional.df <- filter(df,!grepl(pattern="DF",x=SEGNUMBER))
write.table(x=additional.df,file = txt.dir %&% "first-round-variants-additional.txt",
            sep="\t",quote=FALSE,col.names=FALSE,row.names=FALSE)

```


# Manually Adjust Bed File Based on CapSequm Error Report 


Ran CapSequm Stage 1 and downloaded Excel file of errors; manually pasted into txt file and scp to txt directory on server 

Will now read in that info into a data frame and make manual adjustments to the bed file to allow erroneous variants to still be included in experiment  

```{r}

error.df <- fread(txt.dir%&%"CapC_first-round-errors.txt")

small.df <- filter(error.df,grepl("small",Error.detected)) # Will offset by 50 bp 
# PTGFRN,4DF,credible.variant,chr1:117532790 
# ZMIZ1,88DF,credible.variant,chr10:80945964

redun.df <- filter(error.df,grepl("redundant",Error.detected)) # Will need to assign to corresponding oligos after design step

cut.df <- filter(error.df,grepl("cutsite",Error.detected)) # Will offset by 10 bp 
# KLHL42,103DF,credible.variant,chr12:27964781 

df2 <- bed.df

index <- match("PTGFRN,4DF,credible.variant,chr1:117532790",df2$NAME)
df2[index,]$POS0 <- df2[index,]$POS0 + 100 
df2[index,]$POS1 <- df2[index,]$POS1 + 100 
df2[index,]$NAME <- df2[index,]$NAME %&% ".offset100"

index <- match("ZMIZ1,88DF,credible.variant,chr10:80945964",df2$NAME)
df2[index,]$POS0 <- df2[index,]$POS0 + 100 
df2[index,]$POS1 <- df2[index,]$POS1 + 100 
df2[index,]$NAME <- df2[index,]$NAME %&% ".offset100"

index <- match("KLHL42,103DF,credible.variant,chr12:27964781",df2$NAME)
df2[index,]$POS0 <- df2[index,]$POS0 + 10 
df2[index,]$POS1 <- df2[index,]$POS1 + 10 
df2[index,]$NAME <- df2[index,]$NAME %&% ".offset10"


write.table(x=df2,file = txt.dir %&% "first-round-variants_error-adjust.bed",
            sep="\t",quote=FALSE,col.names=FALSE,row.names=FALSE)

```



